### 2021-10-03 project_2.Rmd
Attaching R notebook

---
title: "Project 2"
authors: "Navdeep Singh, Shreya Khobragade, Samridhi Hooda"
output: html_notebook
---

This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

```{r}
plot(cars)
```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.

### 2021-10-03 project_2.Rmd
Adding project's code

@@
 
-This is an [R Markdown](http://rmarkdown.rstudio.com) Notebook. When you execute code within the notebook, the results appear beneath the code.
+```{r load_libraries}
+#loading all the necessary libraries
+library(gutenbergr)
+library(tidytext)
+library(tidyverse)
 
-Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.
-
-```{r}
-plot(cars)
 ```
 
-Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.
+```{r gutenberg_fitzgerald_works}
+#finding texts by Fitzgerald on gutenberg
+fitzgerald_text <- gutenberg_works(author == "Fitzgerald, F. Scott (Francis Scott)")
+```
 
-When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).
-
-The preview shows you a rendered HTML copy of the contents of the editor.
-Consequently, unlike *Knit*, *Preview* does not run any R code chunks.
-Instead, the output of the chunk when it was last run in the editor is displayed.
+```{r gutenberg_fitzgerald_download}
+#downloading all the texts found
+fitzgerald_download <- fitzgerald_text %>% 
+                      gutenberg_download(meta_fields = c("title", "author"))
+```

### 2021-10-06 project_2.Rmd
All texts imported

@@
 library(tidyverse)
-
+library(vroom)
+library(readtext)
 ```
@@
 ```
+
+```{r import_Hemingway_books}
+
+The_Sun_Also_Rises_Hemingway_Ernest <- read_delim("The Sun Also Rises_Hemingway, Ernest.txt", 
+    delim = "\t", quote = "\\\"", escape_double = FALSE, 
+    trim_ws = TRUE)
+The_Old_Man_and_the_Sea_Hemingway_Ernest <- read_delim("The Old Man and the Sea_Hemingway, Ernest.txt", 
+    delim = ".", quote = "\\\"", escape_double = FALSE)
+For_Whom_the_Bell_Tolls_Hemingway_Ernest <- read_delim("For Whom the Bell Tolls_Hemingway, Ernest.txt", 
+    delim = ".", quote = "\\\"", escape_double = FALSE)
+A_Farewell_to_Arms_Hemingway_Ernest <- read_delim("A Farewell to Arms_Hemingway, Ernest.txt", 
+    delim = ".", quote = "\\\"", escape_double = FALSE)
+```
+
+```{r import_Hemingway_bulk}
+#DO NOT RUN
+file_list <- list.files(path = ".", recursive = FALSE, pattern = "*Ernest.txt$", full.names = FALSE)
+
+Hemingway_files <- vroom(file_list, delim = ".", col_names = FALSE)
+```
+
+```{r hemingway_corpus}
+#DO NOT RUN
+DATA_DIR <- "."
+hemingway_corpus <- readtext("A Farewell to Arms_Hemingway, Ernest.txt", docvarsfrom = "filenames", docvarnames = c("title", "author"), devsep = "_")
+```

### 2021-10-06 project_2.Rmd
data modelling

--- 2021-10-03 project_2.Rmd
@@
 library(tidyverse)
-
+library(readtext)
 ```
@@
 ```
+
+```{r import_Hemingway_books}
+
+The_Sun_Also_Rises_Hemingway_Ernest <- read_delim("The Sun Also Rises_Hemingway, Ernest.txt", 
+    delim = "\t", quote = "\\\"", escape_double = FALSE, 
+    trim_ws = TRUE)
+The_Sun_Also_Rises_Hemingway_Ernest <- The_Sun_Also_Rises_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "The Sun Also Rises", .before = text)
+
+The_Old_Man_and_the_Sea_Hemingway_Ernest <- read_delim("The Old Man and the Sea_Hemingway, Ernest.txt", 
+    delim = ".", quote = "\\\"", escape_double = FALSE)
+The_Old_Man_and_the_Sea_Hemingway_Ernest <- The_Old_Man_and_the_Sea_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "The Old Man and the Sea", .before = text)
+
+For_Whom_the_Bell_Tolls_Hemingway_Ernest <- read_delim("For Whom the Bell Tolls_Hemingway, Ernest.txt", 
+    delim = ".", quote = "\\\"", escape_double = FALSE)
+For_Whom_the_Bell_Tolls_Hemingway_Ernest <- For_Whom_the_Bell_Tolls_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "For Whom The Bell Tolls", .before = text)
+
+A_Farewell_to_Arms_Hemingway_Ernest <- read_delim("A Farewell to Arms_Hemingway, Ernest.txt", 
+    delim = ".", quote = "\\\"", escape_double = FALSE)
+A_Farewell_to_Arms_Hemingway_Ernest <- A_Farewell_to_Arms_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "A Farewell to Arms", .before = text)
+
+hemingway_corpus <- bind_rows(The_Sun_Also_Rises_Hemingway_Ernest,The_Old_Man_and_the_Sea_Hemingway_Ernest,For_Whom_the_Bell_Tolls_Hemingway_Ernest,A_Farewell_to_Arms_Hemingway_Ernest)
+```
+
+```{r tidy_hemingway}
+#breaking up texts into words
+tidy_hemingway <- hemingway_corpus %>% unnest_tokens(word, text) %>% anti_join(stop_words)
+```
+
+```{r tidy_fitzgerald}
+#breaking up texts into words
+tidy_fitzgerald <- fitzgerald_download %>% unnest_tokens(word, text) %>% anti_join(stop_words)
+```
+
+```{r adding_female_words}
+#collection of words representing a female
+female_words <- c("she", "her", "woman", "girl", "aunt", "mother", "sister", "wife", "lady")
+
+female_df <- tibble(word = female_words, female = TRUE)
+```
+
+```{r hemingway_tagged}
+hemingway_tagged <- tidy_hemingway %>% left_join(female_df)
+```
+
+```{r fitzgerald_tagged}
+fitzgerald_tagged <- tidy_fitzgerald %>% left_join(female_df)
+```
+
+```{r fitzgerald_count}
+#counting how often female words appeared
+#table shows each text and the respective percentage
+fitzgerald_table <- fitzgerald_tagged %>% group_by(title) %>% count(female) %>% mutate(percent = n/sum(n)*100)
+
+```
+
+```{r hemingway_count}
+#counting how often female words appeared
+#table shows each text and the respective percentage
+hemingway_table <- hemingway_tagged %>% group_by(title) %>% count(female) %>% mutate(percent = n/sum(n)*100)
+
+```
+
+```{r combined_result}
+#table showing two authors and their respective usage of female words
+author_table <- bind_rows(fitzgerald_tagged, hemingway_tagged)
+author_table <- author_table %>% group_by(author) %>% count(female) %>% mutate(percent = n/sum(n)*100) %>% drop_na(female)
+
+#table showing all texts and their respective usage of female words
+combined_table <- bind_rows(hemingway_table, fitzgerald_table) %>% drop_na(female)
+```
+```{r frequency_distribution}
+hemingway_wordcount <- tidy_hemingway %>% count(word, sort = TRUE) %>% mutate(frequency = (n/sum(n)), .after = n)
+fitzgerald_wordcount <- tidy_fitzgerald %>% count(word, sort = TRUE) %>% mutate(frequency = (n/sum(n)), .after = n)
+
+```
+
+```{r exporting_tables}
+save(author_table, file = "saved_data.rda")
+```

### 2021-10-06 project_2.Rmd
Edited Fitzgerald corpus

@@
 fitzgerald_text <- gutenberg_works(author == "Fitzgerald, F. Scott (Francis Scott)")
+fitzgerald_text <- fitzgerald_text[-c(2,3), ]
 ```
@@
 ```
+```{r import_Fitzgerald_books}
+The_Great_Gatsby <- read_delim("The Great Gatsby_Fitzgerald, F. Scott (Francis Scott).txt", 
+    delim = ".", quote = "\\\"", escape_double = FALSE)
+The_Great_Gatsby <- The_Great_Gatsby %>% mutate(author = "Fitzgerald, F. Scott (Francis Scott)", .after = text) %>% mutate(title = "The Great Gatsby", .before = author)
+
+Tender_is_the_Night <- read_delim("Tender is the Night_Fitzgerald, F. Scott (Francis Scott).txt", 
+    delim = ".", quote = "\\\"", escape_double = FALSE)
+Tender_is_the_Night <- Tender_is_the_Night %>% mutate(author = "Fitzgerald, F. Scott (Francis Scott)", .after = text) %>% mutate(title = "Tender is the Night", .before = author)
+```
+
+```{r create_Fitzgerald_corpus}
+fitzgerald_download$gutenberg_id <- NULL
+fitzgerald_corpus <- bind_rows(fitzgerald_download, The_Great_Gatsby, Tender_is_the_Night)
+```
 
@@
     trim_ws = TRUE)
-The_Sun_Also_Rises_Hemingway_Ernest <- The_Sun_Also_Rises_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "The Sun Also Rises", .before = text)
+The_Sun_Also_Rises_Hemingway_Ernest <- The_Sun_Also_Rises_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "The Sun Also Rises", .before = author)
 
@@
     delim = ".", quote = "\\\"", escape_double = FALSE)
-The_Old_Man_and_the_Sea_Hemingway_Ernest <- The_Old_Man_and_the_Sea_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "The Old Man and the Sea", .before = text)
+The_Old_Man_and_the_Sea_Hemingway_Ernest <- The_Old_Man_and_the_Sea_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "The Old Man and the Sea", .before = author)
 
@@
     delim = ".", quote = "\\\"", escape_double = FALSE)
-For_Whom_the_Bell_Tolls_Hemingway_Ernest <- For_Whom_the_Bell_Tolls_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "For Whom The Bell Tolls", .before = text)
+For_Whom_the_Bell_Tolls_Hemingway_Ernest <- For_Whom_the_Bell_Tolls_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "For Whom The Bell Tolls", .before = author)
 
@@
     delim = ".", quote = "\\\"", escape_double = FALSE)
-A_Farewell_to_Arms_Hemingway_Ernest <- A_Farewell_to_Arms_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "A Farewell to Arms", .before = text)
+A_Farewell_to_Arms_Hemingway_Ernest <- A_Farewell_to_Arms_Hemingway_Ernest %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "A Farewell to Arms", .before = author)
 
@@
 #breaking up texts into words
-tidy_fitzgerald <- fitzgerald_download %>% unnest_tokens(word, text) %>% anti_join(stop_words)
+tidy_fitzgerald <- fitzgerald_corpus %>% unnest_tokens(word, text) %>% anti_join(stop_words)
 ```

### 2021-10-08 project_2.Rmd
Generated emotion words

@@
 ```
+
+```{r wildcard}
+puretext <- bind_rows(tidy_hemingway[, 3], tidy_fitzgerald[, 3])
+purestring <- unlist(puretext)
+sad_1 <- paste(unique(str_subset(purestring, "^anger|^angr|^disgust|^contempt|^revuls|^envy|^envi|^jealous|^exaspera|^frustra|^aggrava|^agitat|^annoy|^grouch|^grump|^irrita|^bitter|^dislike|^feroci|^fury|^furi|^hate|^hostil|^loath|^outrage|^rage|^rag|^resent|^scorn|^spite|^venge|^wrath|^torment|^fear|^horr|^alarm|^fright|^hyster|^mortifi|^panic|^shock|^terror|^nervous|^anxi|^apprehen|^distress|^dread|^tense|^uneasi|^uneasy|^worr|^joy|^cheer|^amuse|^bliss|^delight|^ecsta|^elati|^enjoy|^euphor|^gaiety|^glad|^glee|^happi|^jolli|^jolly|^jovial|^jubil|^satisf|^content|^pleasur|^enthrall|^raptur|^eager|^hope|^optimis|^pride|^proud|^triumph|^relie|^zest|^enthusias|^excit|^exhilara|^thrill|^zeal|^love|^ador|^affec|^attrac|^care|^cari|^compassion|^fond|^like|^liking|^sentiment|^tender|^long|^arous|^desire|^infatuat|^lust|^passion|^sad|^disappoint|^dismay|^displeasur|^alien|^defeat|^deject|^embarrass|^homesick|^humiliat|^insecur|^isola|^insult|^lonel|^neglect|^reject|^depress|^despair|^gloom|^glum|^grie|^hopeless|^melancholi|^miser|^sorrow|^unhappi|^woe|^shame|^guilt|^regret|^remorse|^suffer|^agon|^anguish|^hurt|^pity|^piti|^sympath|^surprise|^amaze|^astonish")), collapse = ' ')
+```

### 2021-10-08 project_2.Rmd
Emotion tables

@@
 
+```{r emotion_words}
+emotion_list <- read_csv("emotion_list.txt")
+emotion_words <- tibble(emotion_list, emotion = TRUE)
+```
+
+```{r hemingway_emotions}
+hemingway_emotions <- tidy_hemingway %>% left_join(emotion_words)
+```
+
+```{r fitzgerald_emotions}
+fitzgerald_emotions <- tidy_fitzgerald %>% left_join(emotion_words)
+```
+
+```{r fitzgerald_emotion_table}
+#counting how often female words appeared
+#table shows each text and the respective percentage
+fitzgerald_emotion_calc <- fitzgerald_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = n/sum(n)*100)
+
+```
+
+```{r hemingway_emotion_table}
+#counting how often female words appeared
+#table shows each text and the respective percentage
+hemingway_emotion_calc <- hemingway_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = n/sum(n)*100)
+
+```
+
 ```{r exporting_tables}

### 2021-10-09 project_2.Rmd
Commit
Sentiment Analysis

@@
 library(readtext)
+library(sentimentr)
 ```
@@
 ```{r import_Fitzgerald_books}
-The_Great_Gatsby <- read_delim("The Great Gatsby_Fitzgerald, F. Scott (Francis Scott).txt", 
+The_Great_Gatsby <- read_delim("The Great Gatsby_Fitzgerald, F. Scott (Francis Scott)_clean.txt", 
     delim = ".", quote = "\\\"", escape_double = FALSE)
@@
 
-Tender_is_the_Night <- read_delim("Tender is the Night_Fitzgerald, F. Scott (Francis Scott).txt", 
+Tender_is_the_Night <- read_delim("Tender is the Night_Fitzgerald, F. Scott (Francis Scott)_clean.txt", 
     delim = ".", quote = "\\\"", escape_double = FALSE)
@@
 
-The_Sun_Also_Rises_Hemingway_Ernest <- read_delim("The Sun Also Rises_Hemingway, Ernest.txt", 
+The_Sun_Also_Rises_Hemingway_Ernest <- read_delim("The Sun Also Rises_Hemingway, Ernest_clean.txt", 
     delim = "\t", quote = "\\\"", escape_double = FALSE, 
@@
 
-The_Old_Man_and_the_Sea_Hemingway_Ernest <- read_delim("The Old Man and the Sea_Hemingway, Ernest.txt", 
+The_Old_Man_and_the_Sea_Hemingway_Ernest <- read_delim("The Old Man and the Sea_Hemingway, Ernest_clean.txt", 
     delim = ".", quote = "\\\"", escape_double = FALSE)
@@
 
-For_Whom_the_Bell_Tolls_Hemingway_Ernest <- read_delim("For Whom the Bell Tolls_Hemingway, Ernest.txt", 
+For_Whom_the_Bell_Tolls_Hemingway_Ernest <- read_delim("For Whom the Bell Tolls_Hemingway, Ernest_clean.txt", 
     delim = ".", quote = "\\\"", escape_double = FALSE)
@@
 
-A_Farewell_to_Arms_Hemingway_Ernest <- read_delim("A Farewell to Arms_Hemingway, Ernest.txt", 
+A_Farewell_to_Arms_Hemingway_Ernest <- read_delim("A Farewell to Arms_Hemingway, Ernest_clean.txt", 
     delim = ".", quote = "\\\"", escape_double = FALSE)
@@
 hemingway_corpus <- bind_rows(The_Sun_Also_Rises_Hemingway_Ernest,The_Old_Man_and_the_Sea_Hemingway_Ernest,For_Whom_the_Bell_Tolls_Hemingway_Ernest,A_Farewell_to_Arms_Hemingway_Ernest)
+```
+
+```{r new_sentiment_analysis}
+#temp <- hemingway_corpus %>% unnest_regex(sentences, text, pattern = "[.?!]", to_lower = TRUE)
+my_text <- read_file("The Great Gatsby_Fitzgerald, F. Scott (Francis Scott)_clean.txt")
+my_text <- get_sentences(my_text)
+total_sentiment <- sentiment(my_text)
+total <- sum(total_sentiment$sentiment)
+```
+
+```{r all_authors}
+all_authors <- bind_rows(fitzgerald_corpus, hemingway_corpus)
+
+```
+
+```{r tidy_all_authors}
+tidy_all_authors <- all_authors %>% unnest_tokens(word, text) %>% anti_join(stop_words)
 ```
@@
 
+## Sentiment Analysis
+
+```{r all_author_sentences}
+all_authors_sentence <- all_authors %>%
+  group_by(author, title, text) %>%
+  summarise (text = paste(text, collapse = "")) %>%
+  unnest_regex(sentences, text, pattern = "[.?!]", to_lower = TRUE)  
+```
+
+```{r author_sentence_nr}
+author_sentence_nr <- all_authors_sentence %>%
+  ungroup() %>%
+  mutate(sentence_number = row_number()) %>%
+  group_by(author, title, sentence_number) %>%
+  unnest_tokens(word, sentences) %>%
+  anti_join(stop_words)
+```
+
+```{r author_concept}
+author_concept <-  author_sentence_nr #%>%
+  #left_join(emotion_words) 
+```
+
+```{r author_concept_sentiment}
+author_concept_sentiment <- author_concept %>%
+  inner_join(get_sentiments("bing"))
+```
+
+```{r author_sentiment_total}
+author_sentiment_total <- author_concept_sentiment %>%
+  count(sentence_number, sentiment) %>%
+  pivot_wider(names_from = sentiment,
+              values_from = n,
+              values_fill = 0)  %>%
+  mutate(sentiment = positive - negative) #%>%
+  #left_join(author_concept) #%>%
+  #filter(emotion == TRUE)
+```
+
+```{r author_sentiment_table}
+author_sentiment_table <- author_sentiment_total %>%
+  #pivot_longer(sentiment,
+               #names_to = "word",
+               #values_to = "total_sentiment")  %>%
+  #drop_na() %>%
+  group_by(title, author) %>%
+  summarise (total = sum(sentiment)) %>%
+  ungroup()
+```
+
+```{r author_sentiment_plot}
+author_sentiment_table %>%
+  ggplot(aes(title, y = total, fill = title)) +
+  geom_col(color = "black",
+           alpha = .7,
+           position = "identity") +
+  #facet_wrap(~ concept) +
+  labs(title = "Positive and Negative Sentiment by Concept",
+       x = "Overal Sentiment",
+       y = "Author",
+       fill = "Title") +
+  coord_flip() 
+```
+
 ```{r exporting_tables}
-save(author_table, file = "saved_data.rda")
+save(author_table, hemingway_corpus, file = "saved_data.rda")
 ```

### 2021-10-09 project_2.Rmd
plot

@@
 #table shows each text and the respective percentage
-fitzgerald_emotion_calc <- fitzgerald_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = n/sum(n)*100)
+fitzgerald_emotion_calc <- fitzgerald_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = n/sum(n)*100) %>% mutate(author = "F. Scott Fitzgerald")
 
@@
 #table shows each text and the respective percentage
-hemingway_emotion_calc <- hemingway_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = n/sum(n)*100)
+hemingway_emotion_calc <- hemingway_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = n/sum(n)*100) %>% mutate(author = "Earnest Hemingway")
 
+```
+
+```{r emotion_plot}
+combined_table <- bind_rows(hemingway_emotion_calc, fitzgerald_emotion_calc) %>% filter(emotion == TRUE)
+
+p<-ggplot(data=combined_table, aes(x=percent, y=title, fill=author)) +
+  geom_bar(stat="identity")
+p
 ```

### 2021-10-11 project_2.Rmd
NRC implemented

@@
 library(sentimentr)
+library(ggiraphExtra)
+library(ggplot2)
+library(RColorBrewer)
+library(scales)
 ```
@@
 
-hemingway_corpus <- bind_rows(The_Sun_Also_Rises_Hemingway_Ernest,The_Old_Man_and_the_Sea_Hemingway_Ernest,For_Whom_the_Bell_Tolls_Hemingway_Ernest,A_Farewell_to_Arms_Hemingway_Ernest)
+The_Garden_of_Eden_Hemingway <- read_delim("The Garden of Eden_Hemingway_clean.txt", 
+    delim = "\t", quote = "\\\"", escape_double = FALSE, 
+    trim_ws = TRUE)
+The_Garden_of_Eden_Hemingway_Ernest <- The_Garden_of_Eden_Hemingway %>% mutate(author = "Hemingway, Earnest", .after = text) %>% mutate(title = "The Garden of Eden", .before = author)
+
+hemingway_corpus <- bind_rows(The_Sun_Also_Rises_Hemingway_Ernest,The_Old_Man_and_the_Sea_Hemingway_Ernest,For_Whom_the_Bell_Tolls_Hemingway_Ernest,A_Farewell_to_Arms_Hemingway_Ernest,The_Garden_of_Eden_Hemingway_Ernest)
 ```
@@
 #temp <- hemingway_corpus %>% unnest_regex(sentences, text, pattern = "[.?!]", to_lower = TRUE)
-my_text <- read_file("The Great Gatsby_Fitzgerald, F. Scott (Francis Scott)_clean.txt")
+my_text <- read_file("~/Digital.Humanities/dh_proj_2/The Garden of Eden_Hemingway_clean.txt")
 my_text <- get_sentences(my_text)
+new_text <- c(my_text)
 total_sentiment <- sentiment(my_text)
@@
 #breaking up texts into words
+tidy_eden <- The_Garden_of_Eden_Hemingway_Ernest %>% unnest_tokens(word, text) %>% anti_join(stop_words)
 tidy_hemingway <- hemingway_corpus %>% unnest_tokens(word, text) %>% anti_join(stop_words)
@@
 #breaking up texts into words
+tidy_gatsby <- The_Great_Gatsby %>% unnest_tokens(word, text) %>% anti_join(stop_words)
 tidy_fitzgerald <- fitzgerald_corpus %>% unnest_tokens(word, text) %>% anti_join(stop_words)
+```
+
+```{r sentiment_nrc}
+nrc_list <- get_sentiments("nrc")
+counter <- tidy_hemingway %>% inner_join(nrc_list) %>% count(word, sort = TRUE)
+sentiment_h <- tidy_hemingway %>% inner_join(get_sentiments("nrc") %>% 
+                 filter(sentiment %in% c("anger", 
+                                         "anticipation", "disgust", "fear", "joy", "sadness", "surprise", "trust"))
+    ) %>%
+  count(sentiment) %>%
+  pivot_wider(names_from = sentiment,
+              values_from = n,
+              values_fill = 0) %>% mutate(author = "Hemingway")#%>% 
+  #mutate(sentiment = positive - negative)
+sentiment_h3 <- tidy_eden %>% inner_join(get_sentiments("nrc") %>% 
+                 filter(sentiment %in% c("anger", 
+                                         "anticipation", "disgust", "fear", "joy", "sadness", "surprise", "trust"))
+    ) %>%
+  count(sentiment) %>%
+  pivot_wider(names_from = sentiment,
+              values_from = n,
+              values_fill = 0) %>% mutate(author = "Hemingway")
+sentiment_h2 <- sentiment_h[,-9]/126688
+sentiment_h2 <- sentiment_h2*100
+sentiment_h2 <- sentiment_h2 %>% mutate(author = "Hemingway")
+#anger	anticipation	disgust	fear	joy	sadness	surprise	trust
+```
+
+```{r}
+nrc_joy <- get_sentiments("nrc")
+#tidy_hemingway %>% inner_join(nrc_list) %>% count(word, sort = TRUE)
+sentiment_f <- tidy_fitzgerald %>% inner_join(get_sentiments("nrc") %>% 
+                 filter(sentiment %in% c("anger", 
+                                         "anticipation", "disgust", "fear", "joy", "sadness", "surprise", "trust"))
+    ) %>%
+  count(sentiment) %>%
+  pivot_wider(names_from = sentiment,
+              values_from = n,
+              values_fill = 0) %>% mutate(author = "Fitzgerald")
+sentiment_f3 <- tidy_gatsby %>% inner_join(get_sentiments("nrc") %>% 
+                 filter(sentiment %in% c("anger", 
+                                         "anticipation", "disgust", "fear", "joy", "sadness", "surprise", "trust"))
+    ) %>%
+  count(sentiment) %>%
+  pivot_wider(names_from = sentiment,
+              values_from = n,
+              values_fill = 0) %>% mutate(author = "Fitzgerald")
+sentiment_f2 <- sentiment_f[,-9]/134829
+sentiment_f2 <- sentiment_f2*100
+sentiment_f2 <- sentiment_f2 %>% mutate(author = "Fitzgerald")
+```
+
+```{r}
+sentiment_table <- bind_rows(sentiment_h, sentiment_f)
+#sentiment_table2 <- bind_rows(sentiment_f3, sentiment_h3)
+ggRadar(
+  data = sentiment_table,
+  mapping = aes(color = author),
+  rescale = FALSE,
+  interactive = TRUE,
+  use.label = TRUE,
+  size = 2,
+  legend.position = "right"
+)
+```
+```{r}
+ggplot(data=sentiment_table, aes(x=anger, y=anticipation, fill=author)) +
+  geom_bar(stat="identity")
+
 ```

### 2021-10-12 project_2.Rmd
clean up code

--- 2021-10-09 project_2.Rmd
@@
 
-```{r adding_female_words}
-#collection of words representing a female
-female_words <- c("she", "her", "woman", "girl", "aunt", "mother", "sister", "wife", "lady")
-
-female_df <- tibble(word = female_words, female = TRUE)
+```{r sentiment_nrc}
+nrc_list <- get_sentiments("nrc")
+counter <- tidy_hemingway %>% inner_join(nrc_list) %>% count(word, sort = TRUE)
+sentiment_abs_h <- tidy_hemingway %>% inner_join(get_sentiments("nrc") %>% 
+                 filter(sentiment %in% c("anger", 
+                                         "anticipation", "disgust", "fear", "joy", "sadness", "surprise", "trust"))
+    ) %>%
+  count(sentiment) %>%
+  pivot_wider(names_from = sentiment,
+              values_from = n,
+              values_fill = 0) %>% mutate(author = "Hemingway")#%>% 
+  #mutate(sentiment = positive - negative)
+total_words_hemingway <- NROW(tidy_hemingway)
+sentiment_rel_h <- sentiment_abs_h[,-9]/total_words_hemingway
+sentiment_rel_h <- sentiment_rel_h*100
+sentiment_rel_h <- sentiment_rel_h %>% mutate(author = "Hemingway")
+#anger	anticipation	disgust	fear	joy	sadness	surprise	trust
 ```
 
-```{r hemingway_tagged}
-hemingway_tagged <- tidy_hemingway %>% left_join(female_df)
+```{r}
+nrc_joy <- get_sentiments("nrc")
+#tidy_hemingway %>% inner_join(nrc_list) %>% count(word, sort = TRUE)
+sentiment_abs_f <- tidy_fitzgerald %>% inner_join(get_sentiments("nrc") %>% 
+                 filter(sentiment %in% c("anger", 
+                                         "anticipation", "disgust", "fear", "joy", "sadness", "surprise", "trust"))
+    ) %>%
+  count(sentiment) %>%
+  pivot_wider(names_from = sentiment,
+              values_from = n,
+              values_fill = 0) %>% mutate(author = "Fitzgerald")
+
+total_words_fitzgerald <- NROW(tidy_fitzgerald)
+sentiment_rel_f <- sentiment_abs_f[,-9]/total_words_fitzgerald
+sentiment_rel_f <- sentiment_rel_f*100
+sentiment_rel_f <- sentiment_rel_f %>% mutate(author = "Fitzgerald")
 ```
 
-```{r fitzgerald_tagged}
-fitzgerald_tagged <- tidy_fitzgerald %>% left_join(female_df)
-```
+```{r}
+sentiment_both <- bind_rows(sentiment_rel_h, sentiment_rel_f)
 
-```{r fitzgerald_count}
-#counting how often female words appeared
-#table shows each text and the respective percentage
-fitzgerald_table <- fitzgerald_tagged %>% group_by(title) %>% count(female) %>% mutate(percent = n/sum(n)*100)
-
-```
-
-```{r hemingway_count}
-#counting how often female words appeared
-#table shows each text and the respective percentage
-hemingway_table <- hemingway_tagged %>% group_by(title) %>% count(female) %>% mutate(percent = n/sum(n)*100)
-
-```
-
-```{r combined_result}
-#table showing two authors and their respective usage of female words
-author_table <- bind_rows(fitzgerald_tagged, hemingway_tagged)
-author_table <- author_table %>% group_by(author) %>% count(female) %>% mutate(percent = n/sum(n)*100) %>% drop_na(female)
-
-#table showing all texts and their respective usage of female words
-combined_table <- bind_rows(hemingway_table, fitzgerald_table) %>% drop_na(female)
-```
-```{r frequency_distribution}
-hemingway_wordcount <- tidy_hemingway %>% count(word, sort = TRUE) %>% mutate(frequency = (n/sum(n)), .after = n)
-fitzgerald_wordcount <- tidy_fitzgerald %>% count(word, sort = TRUE) %>% mutate(frequency = (n/sum(n)), .after = n)
-
+ggRadar(
+  data = sentiment_both,
+  mapping = aes(color = author),
+  rescale = FALSE,
+  interactive = TRUE,
+  use.label = TRUE,
+  size = 2,
+  legend.position = "right"
+)
 ```
@@
 
-## Sentiment Analysis
-
-```{r all_author_sentences}
-all_authors_sentence <- all_authors %>%
-  group_by(author, title, text) %>%
-  summarise (text = paste(text, collapse = "")) %>%
-  unnest_regex(sentences, text, pattern = "[.?!]", to_lower = TRUE)  
-```
-
-```{r author_sentence_nr}
-author_sentence_nr <- all_authors_sentence %>%
-  ungroup() %>%
-  mutate(sentence_number = row_number()) %>%
-  group_by(author, title, sentence_number) %>%
-  unnest_tokens(word, sentences) %>%
-  anti_join(stop_words)
-```
-
-```{r author_concept}
-author_concept <-  author_sentence_nr #%>%
-  #left_join(emotion_words) 
-```
-
-```{r author_concept_sentiment}
-author_concept_sentiment <- author_concept %>%
-  inner_join(get_sentiments("bing"))
-```
-
-```{r author_sentiment_total}
-author_sentiment_total <- author_concept_sentiment %>%
-  count(sentence_number, sentiment) %>%
-  pivot_wider(names_from = sentiment,
-              values_from = n,
-              values_fill = 0)  %>%
-  mutate(sentiment = positive - negative) #%>%
-  #left_join(author_concept) #%>%
-  #filter(emotion == TRUE)
-```
-
-```{r author_sentiment_table}
-author_sentiment_table <- author_sentiment_total %>%
-  #pivot_longer(sentiment,
-               #names_to = "word",
-               #values_to = "total_sentiment")  %>%
-  #drop_na() %>%
-  group_by(title, author) %>%
-  summarise (total = sum(sentiment)) %>%
-  ungroup()
-```
-
-```{r author_sentiment_plot}
-author_sentiment_table %>%
-  ggplot(aes(title, y = total, fill = title)) +
-  geom_col(color = "black",
-           alpha = .7,
-           position = "identity") +
-  #facet_wrap(~ concept) +
-  labs(title = "Positive and Negative Sentiment by Concept",
-       x = "Overal Sentiment",
-       y = "Author",
-       fill = "Title") +
-  coord_flip() 
-```
-
 ```{r exporting_tables}
-save(author_table, hemingway_corpus, file = "saved_data.rda")
+save(author_table, hemingway_corpus, fitzgerald_corpus, file = "saved_data.rda")
 ```

### 2021-10-12 NER/NER_analysis.Rmd
NER Analysis

---
title: "NLP and NER"
author: "Navdeep Singh"
date: "13/10/2021"
output: html_document
---

# Part 1: Initial setup

## Installing NLP libraries

```{r error=FALSE, eval= FALSE, warning=FALSE, message=FALSE}
# Run only once
install.packages("rJava")
install.packages("openNLP")
install.packages("NLP")
install.packages("tcltk2")
```

## Downloading NLP models

```{r echo=TRUE, eval= FALSE, error=FALSE, warning=FALSE, message=FALSE}
# Run only once
install.packages("openNLPmodels.en", dependencies=TRUE, repos = "http://datacube.wu.ac.at/")
```

## Importing all libraries

```{r echo=TRUE, results='hide', message=FALSE}
#NLP Libraries
library(rJava)
library(openNLP)
library(NLP)

#Tidy data manipulation
library(stringr)
library(dplyr)
library(tidyr)
library(tidytext)
library(readr)
library(stringi)
library(textclean)

#Ingest data
load("saved_data.rda")

#Helper library
library(sqldf)
library(tcltk2)

#Graphics library
library(ggiraphExtra)
library(ggplot2)
library(RColorBrewer)
library(scales)
```

# Part 2: Ingesting corpus

## Importing corpus
```{r import_corpus}
corpus <- bind_rows(hemingway_corpus, fitzgerald_corpus)
```

## Cleaning data

### Stripping extraneous characters

```{r pre_clean_corpus}
corpus_clean <- corpus %>%
  filter(text != "") %>%
  mutate(text = str_replace_all(text, "_", " ")) %>%
  mutate(text = replace_contraction(text)) %>%
  mutate(text = replace_curly_quote(text))
```

## Part 3: Preparing data for NER

### Collapsing content of novels into a string

```{r create_strings}
corpus_text <- corpus_clean %>%
  group_by(title) %>%
  mutate(text = paste(as.character(text), collapse = " ")) %>%
  distinct() %>%
  ungroup()
```

### Converting to nested String object

```{r create_nested_string}
corpus_text_str <- corpus_text %>%
  group_by(title) %>%
  mutate(text = list(as.String(text)))
```

### Setting up NLP pipeline

#### Setting up annotation

```{r initiate_pipeline}
#set pipeline
wordAnnotator <- Maxent_Word_Token_Annotator(language = "en")
sentenceAnnotator <- Maxent_Sent_Token_Annotator(language = "en")
characterAnnotatorEN <- Maxent_Entity_Annotator(language = "en", kind = "person")

pipeline <- list(sentenceAnnotator, wordAnnotator, characterAnnotatorEN, locationAnnotatorEN)
```

### Initiating annotation

#### Chunking and extracting persons

```{r NER_chunker, message=FALSE}
#create empty df
full_df = as.data.frame(NULL)
chunk_size = 10000

for (j in 1:nrow(corpus_text_str)) {
  #get number of chunks
  chunk <- nchar(corpus_text_str$text[j]) %/% chunk_size
  text <- unlist(corpus_text_str$text[j])
  text <- as.String(text)

  #Loop runs through the text section by section and reads each chunk into a df

  for (i in 1:chunk) {
    print(paste0(
      "Processing title: ",
      corpus_text_str$title[j],
      " - section ",
      i,
      " of ",
      chunk
    ))
    temp_df = NULL

    if (i == 1) {
      m = 1
    }

    if (i == chunk) {
      m = n + 1
      n = (nchar(text))
    }
    else{
      n <- m + chunk_size
    }

    temp_string = text[m, n]

    temp_ann <- NLP::annotate(temp_string, pipeline)

    temp_df <-  temp_ann %>%
      as.data.frame %>% 
      filter(type != "word")

    temp_df <- temp_df %>%
      mutate(words = str_sub(
        as.character(temp_string),
        start = temp_df$start,
        end = temp_df$end
      )) %>%
      unnest_wider(features)

    temp_df <- temp_df %>%
      mutate(author = corpus_text_str$author[j], title = corpus_text_str$title[j])

    # Stitching together all data
    full_df <- full_df %>%
      bind_rows(temp_df)

    m <- m + chunk_size
  }
}

```

#### Creating backup variable

```{r backup_annotated_data}
full_df_backup <- full_df
```

#### Rough cleaning the data

```{r rough_clean}
full_df <-  full_df %>%
  mutate(words = str_remove_all(words, '[:punct:]'))
```

#### Realigning the columns

```{r reshape_df}
full_df <- full_df %>%
  relocate(c("author", "title"), .before = 1) %>%
  select(-id,-constituents) 
```

#### Backing up processed data

```{r backup_annotations, results='hide'}
write_csv(full_df, "annotation_backup.csv") 
```

#### Restoring data from backup (Optional)

```{r restore_annotated_data}
#full_df <- read_csv("annotation_backup.csv")
```

## Part 4: Cleaning annotated text

### Appending locations in each sentence

```{r split_sentence_entity}
df1 <- full_df %>%
  filter(type == "sentence") %>%
  mutate(sentence_num = row_number()) %>%
  select(author, title, words, sentence_num) %>%
  rename(sentence = words)

df2 <-  full_df %>%
  filter(type == "entity") %>%
  mutate(record = row_number()) %>%
  select(title, words, kind)
```

### Cleaning entities thoroughly

```{r computational_cleaning}
df2 <- df2 %>%
  mutate(words = str_replace_all(words, "Dear ", "")) %>%
  mutate(words = str_replace_all(words, "Young ", "")) %>%
  mutate(words = str_replace_all(words, "Ah", "")) %>%
  mutate(words = str_replace_all(words, "Oh", "")) %>%
  mutate(words = str_trim(words, side = "both")) %>%
  mutate(words = str_trim(gsub("[A-Z]{2,}", "", words))) %>%
  mutate(words = str_squish(words)) %>%
  mutate_all( ~ ifelse(. %in% c("N/A", "null", ""), NA, .)) %>%
  drop_na() %>%
  dplyr::filter(nchar(words) > 2) %>%
  distinct()

capital_stopwords <-
  as.data.frame(str_to_title(stop_words$word)) %>%
  rename(words = 1)

df2 <- df2 %>%
  anti_join(capital_stopwords)
```

### Manually cleaning entities

#### Exporting the table of entities

```{r backup_join}
write_csv(df2, "pre_join_entities.csv")
```

#### Importing the cleaned table of entities

```{r load_cleaned_data}
#Prep the data
pre_join <- read_csv("pre_join_entities_clean.csv", na = "NA")
```

#### Removing duplicates

```{r limit_join_words}
pre_join <- pre_join %>%
  select(words, kind) %>%
  drop_na()  %>%
  distinct()
```

### Joining entities to sentences

```{r match_dataframes}
#Execute a SQL query
full_join_df <- sqldf("SELECT *
      FROM df1
      LEFT JOIN pre_join ON df1.sentence LIKE '%' || pre_join.words || '%'")

full_join_df <- full_join_df %>%
  distinct()
```

## Part 5: Sentiment analysis

### Cleaning entities + sentences table

```{r exporting_entities_raw}
write_csv(full_join_df, "entities_raw.csv")
```

```{r importing_entities_clean}
clean_entities <- read.csv("entities_cleaned.csv")
```

### Attaching sentiments to entities

#### Unnesting sentences

```{r unnest}
entities_unnest <- clean_entities %>%
  unnest_tokens(word, sentence)
```

#### Dropping Stopwords

```{r remove_stopwords}
entities_unnest <- entities_unnest %>%
  anti_join(stop_words)
```

#### Attaching NRC sentiments

```{r message=FALSE}
#create sentiment table
entities_sentiment <- entities_unnest %>%
  group_by(author, title) %>%
  inner_join(get_sentiments("nrc")) %>%
  count(sentence_nr, sentiment) %>%
  spread(sentiment, n, fill = 0) %>%
  mutate(sentiment = positive - negative)
```

#### Combining sentiments with entitiees

```{r message=FALSE}
entities_matches_sentiment <- entities_unnest %>%
  inner_join(entities_sentiment) %>%
  distinct_at(vars(-word))  
```

#### Creating sentiment table

```{r message=FALSE}
ner_total_sentiment <- entities_matches_sentiment %>%
  group_by(author, title, words, kind) %>%
  summarise(total = mean(sentiment))  
```

## Part 6: Visualization

### Chart of the top 10 positive entities

```{r message=FALSE}
ner_total_sentiment %>%
  group_by(title) %>%
  filter(kind == "person") %>%
  top_n(2) %>%
  mutate(words = reorder(words, total)) %>%
  ggplot(aes(words, y = total, fill = title)) +
  geom_col() +
  facet_wrap(~ title, scales = "free") +
  coord_flip()
```

### Chart of the top 10 negative entities

```{r message=FALSE}
ner_total_sentiment %>%
  group_by(title) %>%
  filter(kind == "person") %>%
  top_n(-5) %>%
  mutate(words = reorder(words, (desc(total)))) %>%
  ggplot(aes(words, y = total, fill = title)) +
  geom_col() +
  facet_wrap(~ title, scales = "free") +
  coord_flip()
```

### Visualising entities with NRC emotions

#### Ranking by "total" emotion

```{r radar_plot_total, message=FALSE}
radar_facet <- entities_matches_sentiment %>%
  select(-positive, -negative, -sentiment) %>% #drop out the unnecessary columns 
  group_by(title, words, kind) %>%
  summarise(across(anger:trust, sum)) %>%
  mutate(total = rowSums(across(where(is.numeric))))  %>%
  arrange(desc(total)) %>%
  head(10)  %>% #Change number to include more or fewer entities
  mutate(across(anger:trust, .fns = ~ round((. / total) * 100))) %>%
  select(-total,-kind)

ggRadar(
  data = radar_facet,
  mapping = aes(color = title, facet = words),
  rescale = FALSE,
  interactive = TRUE,
  use.label = TRUE,
  size = 2,
  legend.position = "right"
)
```

#### Radar plot by the highest positive or negative emotion

```{r message=FALSE}
radar_facet_sentiment <- entities_matches_sentiment %>%
  #Change filter to locations for locations
  filter(words == "Dick") %>%
  group_by(title, words, kind) %>%
  summarise(across(anger:sentiment, sum)) %>%
  arrange(desc(sentiment))  %>%
  head(5)  %>% #Change number to include more or fewer entities
  select(-positive, -negative, -sentiment, -kind)

tender_is_the_night_protagonist <- entities_matches_sentiment %>% filter(title == "Tender is the Night") %>%
  #Change filter to locations for locations
  filter(words == "Dick Diver") %>%
  group_by(title, words, kind) %>%
  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)

the_great_gatsby_protagonist <- entities_matches_sentiment %>%
  #Change filter to locations for locations
  filter(words == "Gatsby") %>%
  group_by(title, words, kind) %>%
  summarise(across(anger:sentiment, sum)) %>% filter(title == "The Great Gatsby") %>% select(-positive, -negative, -sentiment, -kind)

tggp <- the_great_gatsby_protagonist %>% filter(title == "The Great Gatsby") %>% summarise(across(anger:trust, sum))

data_for_gatsby <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Gats"), ] %>% filter(title == "The Great Gatsby") %>%
  group_by(title, words, kind) %>%
  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)

data_for_tender <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Dick"), ] %>% filter(title == "Tender is the Night") %>%
  group_by(title, words, kind) %>%
  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
temp_table <- data_for_gatsby %>% ungroup() %>% select(-title, -words)
#sum_anger <- data_for_gatsby %>% sum(anger)
data_for_gatsby_final <- mapply(sum, temp_table)

radar_facet_sentiment_2 <- bind_rows(tender_is_the_night_protagonist, the_great_gatsby_protagonist)

ggRadar(
  data = radar_facet_sentiment_2,
  mapping = aes(color = title, facet = words),
  rescale = FALSE,
  interactive = TRUE,
  use.label = TRUE,
  size = 2,
  legend.position = "right"
)
```

### 2021-10-12 project_2.Rmd
NER Analysis

@@
 hemingway_corpus <- bind_rows(The_Sun_Also_Rises_Hemingway_Ernest,The_Old_Man_and_the_Sea_Hemingway_Ernest,For_Whom_the_Bell_Tolls_Hemingway_Ernest,A_Farewell_to_Arms_Hemingway_Ernest,The_Garden_of_Eden_Hemingway_Ernest)
-```
-
-```{r new_sentiment_analysis}
-#temp <- hemingway_corpus %>% unnest_regex(sentences, text, pattern = "[.?!]", to_lower = TRUE)
-my_text <- read_file("~/Digital.Humanities/dh_proj_2/The Garden of Eden_Hemingway_clean.txt")
-my_text <- get_sentences(my_text)
-new_text <- c(my_text)
-total_sentiment <- sentiment(my_text)
-total <- sum(total_sentiment$sentiment)
 ```

### 2021-10-13 NER_analysis_copy.Rmd
Complete NER Analysis

--- 2021-10-12 NER/NER_analysis.Rmd
@@
 library(sqldf)
-library(tcltk2)
 
@@
 
-pipeline <- list(sentenceAnnotator, wordAnnotator, characterAnnotatorEN, locationAnnotatorEN)
+pipeline <- list(sentenceAnnotator, wordAnnotator, characterAnnotatorEN)
 ```
@@
 ```{r backup_join}
-write_csv(df2, "pre_join_entities.csv")
+write_csv(df2, "pre_join_entities_new.csv")
 ```
@@
 ```{r importing_entities_clean}
-clean_entities <- read.csv("entities_cleaned.csv")
+clean_entities <- read.csv("entities_clean.csv")
 ```
@@
   inner_join(get_sentiments("nrc")) %>%
-  count(sentence_nr, sentiment) %>%
+  count(sentence_num, sentiment) %>%
   spread(sentiment, n, fill = 0) %>%
@@
 ```{r message=FALSE}
-ner_total_sentiment <- entities_matches_sentiment %>%
+ner_total_sentiment <- entities_matches_sentiment %>% drop_na(words) %>%
   group_by(author, title, words, kind) %>%
@@
   filter(kind == "person") %>%
-  top_n(2) %>%
+  top_n(5) %>%
   mutate(words = reorder(words, total)) %>%
@@
 ```{r radar_plot_total, message=FALSE}
-radar_facet <- entities_matches_sentiment %>%
+radar_facet <- entities_matches_sentiment %>% drop_na(words) %>%
   select(-positive, -negative, -sentiment) %>% #drop out the unnecessary columns 
@@
   #Change filter to locations for locations
-  filter(words == "Dick") %>%
+  filter(kind == "person") %>%
   group_by(title, words, kind) %>%
@@
   arrange(desc(sentiment))  %>%
-  head(5)  %>% #Change number to include more or fewer entities
+  head(10)  %>% #Change number to include more or fewer entities
   select(-positive, -negative, -sentiment, -kind)
@@
 
+data_for_gatsby <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, c("Gats", "Jay")), ] %>% filter(title == "The Great Gatsby") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+new_gatsby <- data_for_gatsby[c(1:3),c(3:10)] %>%  summarise(across(anger:trust, sum)) %>% mutate(protagonist = "Gatsby", .before  = anger)
+new_gatsby_sum <- bind_rows(data_for_gatsby, new_gatsby)
+
+data_for_diver <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Dick"), ] %>% filter(title == "Tender is the Night") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+data_for_anthony <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Antho"), ] %>% filter(title == "The Beautiful and Damned") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+data_for_amory <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Amor"), ] %>% filter(title == "This Side of Paradise") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
 data_for_gatsby <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Gats"), ] %>% filter(title == "The Great Gatsby") %>%
@@
 
-data_for_tender <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Dick"), ] %>% filter(title == "Tender is the Night") %>%
+data_for_daisy <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Dais"), ] %>% filter(title == "The Great Gatsby") %>%
   group_by(title, words, kind) %>%
   summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
-temp_table <- data_for_gatsby %>% ungroup() %>% select(-title, -words)
+
+data_for_gatsby_2 <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Jay"), ] %>% filter(title == "The Great Gatsby") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+data_for_henry <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Cath"), ] %>% filter(title == "A Farewell to Arms") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+temp_table1 <-bind_rows(data_for_gatsby, data_for_gatsby_2) %>% ungroup() %>% select(-title, -words)
+temp_table2 <- data_for_diver %>% ungroup() %>% select(-title, -words)
+temp_table3 <- data_for_anthony %>% ungroup() %>% select(-title, -words)
+temp_table4 <- data_for_amory %>% ungroup() %>% select(-title, -words)
+temp_table5 <- data_for_daisy %>% ungroup() %>% select(-title, -words)
+temp_table6 <- data_for_henry %>% ungroup() %>% select(-title, -words)
 #sum_anger <- data_for_gatsby %>% sum(anger)
-data_for_gatsby_final <- mapply(sum, temp_table)
+data_for_gatsby_final <- mapply(sum, temp_table1)
+data_for_diver_final <- mapply(sum, temp_table2)
+data_for_anthony_final <- mapply(sum, temp_table3)
+data_for_amory_final <- mapply(sum, temp_table4)
+data_for_daisy_final <- mapply(sum, temp_table5)
+data_for_henry_final <- mapply(sum, temp_table6)
+
+protag_fitz <- read_csv("protag_f.csv") %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((. / total) * 100)))
+Gatsby_scale <- protag_fitz %>% filter(protagonist == "Gatsby")
 
@@
 ggRadar(
-  data = radar_facet_sentiment_2,
-  mapping = aes(color = title, facet = words),
+  data = protag_fitz %>% select(-total),
+  mapping = aes(color = protagonist, facet = protagonist),
   rescale = FALSE,

### 2021-10-13 NER_analysis_final.Rmd
NER Analysis Completed

--- 2021-10-13 NER_analysis_copy.Rmd
@@
   use.label = TRUE,
-  size = 2,
+  size = 1,
   legend.position = "right"
@@
   head(10)  %>% #Change number to include more or fewer entities
-  select(-positive, -negative, -sentiment, -kind)
+  select(-positive, -negative, -sentiment, -kind) %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((. / total) * 100)))
 
-tender_is_the_night_protagonist <- entities_matches_sentiment %>% filter(title == "Tender is the Night") %>%
+sentiment_Dick_Diver <- entities_matches_sentiment %>% filter(title == "Tender is the Night") %>%
   #Change filter to locations for locations
-  filter(words == "Dick Diver") %>%
+  filter(words == "Dick") %>%
   group_by(title, words, kind) %>%
@@
 
-the_great_gatsby_protagonist <- entities_matches_sentiment %>%
+sentiment_Gatsby <- entities_matches_sentiment %>%
   #Change filter to locations for locations
@@
 
-data_for_anthony <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Antho"), ] %>% filter(title == "The Beautiful and Damned") %>%
+sentiment_Anthony <- entities_matches_sentiment %>% filter(words == "Anthony") %>% filter(title == "The Beautiful and Damned") %>%
   group_by(title, words, kind) %>%
@@
 
-data_for_amory <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Amor"), ] %>% filter(title == "This Side of Paradise") %>%
+sentiment_Amory <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Amor"), ] %>% filter(title == "This Side of Paradise") %>%
   group_by(title, words, kind) %>%
@@
 
-data_for_henry <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Cath"), ] %>% filter(title == "A Farewell to Arms") %>%
+sentiment_Henry <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Henry"), ] %>% filter(title == "A Farewell to Arms") %>%
   group_by(title, words, kind) %>%
   summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+all_protag <- bind_rows(sentiment_Gatsby, sentiment_Anthony, sentiment_Amory, sentiment_Dick_Diver) %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((.
+/ total) * 100)))
 
@@
 ggRadar(
-  data = protag_fitz %>% select(-total),
-  mapping = aes(color = protagonist, facet = protagonist),
+  # Add select(-total)
+  data = all_protag %>% select(-total),
+  mapping = aes(color = title, facet = words),
   rescale = FALSE,
@@
 ```
+
+```{r graphs_hemingway}
+sentiment_Henry <- entities_matches_sentiment %>% filter(title == "A Farewell to Arms") %>%
+  #Change filter to locations for locations
+  filter(words == "Henry") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+sentiment_Robert <- entities_matches_sentiment %>% filter(title == "For Whom The Bell Tolls") %>%
+  #Change filter to locations for locations
+  filter(words == "Robert Jordan") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+sentiment_Jake <- entities_matches_sentiment %>% filter(title == "The Sun Also Rises") %>%
+  #Change filter to locations for locations
+  filter(words == "Barnes") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+sentiment_David <- entities_matches_sentiment %>% filter(title == "The Garden of Eden") %>%
+  #Change filter to locations for locations
+  filter(words == "David") %>%
+  group_by(title, words, kind) %>%
+  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
+
+all_protag_h <- bind_rows(sentiment_Henry, sentiment_Robert, sentiment_Jake, sentiment_David) %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((.
+/ total) * 100)))
+
+ggRadar(
+  # Add select(-total)
+  data = all_protag_h %>% select(-total),
+  mapping = aes(color = title, facet = words),
+  rescale = FALSE,
+  interactive = TRUE,
+  use.label = TRUE,
+  size = 2,
+  legend.position = "right"
+)
+
+```

### 2021-10-13 NER_analysis_final.Rmd
Cleaning up code

@@
 
-```{r message=FALSE}
+```{r message=FALSE graphs_fitzgerald}
 radar_facet_sentiment <- entities_matches_sentiment %>%
-  #Change filter to locations for locations
+
   filter(kind == "person") %>%
@@
 sentiment_Dick_Diver <- entities_matches_sentiment %>% filter(title == "Tender is the Night") %>%
-  #Change filter to locations for locations
   filter(words == "Dick") %>%
@@
 sentiment_Gatsby <- entities_matches_sentiment %>%
-  #Change filter to locations for locations
   filter(words == "Gatsby") %>%
@@
   summarise(across(anger:sentiment, sum)) %>% filter(title == "The Great Gatsby") %>% select(-positive, -negative, -sentiment, -kind)
-
-tggp <- the_great_gatsby_protagonist %>% filter(title == "The Great Gatsby") %>% summarise(across(anger:trust, sum))
-
-data_for_gatsby <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, c("Gats", "Jay")), ] %>% filter(title == "The Great Gatsby") %>%
-  group_by(title, words, kind) %>%
-  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
-
-new_gatsby <- data_for_gatsby[c(1:3),c(3:10)] %>%  summarise(across(anger:trust, sum)) %>% mutate(protagonist = "Gatsby", .before  = anger)
-new_gatsby_sum <- bind_rows(data_for_gatsby, new_gatsby)
-
-data_for_diver <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Dick"), ] %>% filter(title == "Tender is the Night") %>%
-  group_by(title, words, kind) %>%
-  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
 
@@
 
-data_for_gatsby <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Gats"), ] %>% filter(title == "The Great Gatsby") %>%
-  group_by(title, words, kind) %>%
-  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
-
-data_for_daisy <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Dais"), ] %>% filter(title == "The Great Gatsby") %>%
-  group_by(title, words, kind) %>%
-  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
-
-data_for_gatsby_2 <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Jay"), ] %>% filter(title == "The Great Gatsby") %>%
-  group_by(title, words, kind) %>%
-  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
-
 sentiment_Henry <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Henry"), ] %>% filter(title == "A Farewell to Arms") %>%
@@
 
-all_protag <- bind_rows(sentiment_Gatsby, sentiment_Anthony, sentiment_Amory, sentiment_Dick_Diver) %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((.
+all_protag_fitz <- bind_rows(sentiment_Gatsby, sentiment_Anthony, sentiment_Amory, sentiment_Dick_Diver) %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((.
 / total) * 100)))
 
-temp_table1 <-bind_rows(data_for_gatsby, data_for_gatsby_2) %>% ungroup() %>% select(-title, -words)
-temp_table2 <- data_for_diver %>% ungroup() %>% select(-title, -words)
-temp_table3 <- data_for_anthony %>% ungroup() %>% select(-title, -words)
-temp_table4 <- data_for_amory %>% ungroup() %>% select(-title, -words)
-temp_table5 <- data_for_daisy %>% ungroup() %>% select(-title, -words)
-temp_table6 <- data_for_henry %>% ungroup() %>% select(-title, -words)
-#sum_anger <- data_for_gatsby %>% sum(anger)
-data_for_gatsby_final <- mapply(sum, temp_table1)
-data_for_diver_final <- mapply(sum, temp_table2)
-data_for_anthony_final <- mapply(sum, temp_table3)
-data_for_amory_final <- mapply(sum, temp_table4)
-data_for_daisy_final <- mapply(sum, temp_table5)
-data_for_henry_final <- mapply(sum, temp_table6)
+ggRadar(
 
-protag_fitz <- read_csv("protag_f.csv") %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((. / total) * 100)))
-Gatsby_scale <- protag_fitz %>% filter(protagonist == "Gatsby")
-
-radar_facet_sentiment_2 <- bind_rows(tender_is_the_night_protagonist, the_great_gatsby_protagonist)
-
-ggRadar(
-  # Add select(-total)
-  data = all_protag %>% select(-total),
+  data = all_protag_fitz %>% select(-total),
   mapping = aes(color = title, facet = words),
@@
 sentiment_Henry <- entities_matches_sentiment %>% filter(title == "A Farewell to Arms") %>%
-  #Change filter to locations for locations
   filter(words == "Henry") %>%
@@
 sentiment_Robert <- entities_matches_sentiment %>% filter(title == "For Whom The Bell Tolls") %>%
-  #Change filter to locations for locations
   filter(words == "Robert Jordan") %>%
@@
 sentiment_Jake <- entities_matches_sentiment %>% filter(title == "The Sun Also Rises") %>%
-  #Change filter to locations for locations
   filter(words == "Barnes") %>%
@@
 sentiment_David <- entities_matches_sentiment %>% filter(title == "The Garden of Eden") %>%
-  #Change filter to locations for locations
   filter(words == "David") %>%
@@
 
-all_protag_h <- bind_rows(sentiment_Henry, sentiment_Robert, sentiment_Jake, sentiment_David) %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((.
+all_protag_hemi <- bind_rows(sentiment_Henry, sentiment_Robert, sentiment_Jake, sentiment_David) %>% mutate(total = rowSums(across(where(is.numeric)))) %>% mutate(across(anger:trust, .fns = ~ round((.
 / total) * 100)))
@@
 ggRadar(
-  # Add select(-total)
-  data = all_protag_h %>% select(-total),
+
+  data = all_protag_hemi %>% select(-total),
   mapping = aes(color = title, facet = words),

### 2021-10-14 Submission.Rmd
Final HTML

---
title: "Digital Humanities - Ashoka University"
author: "Team members: Navdeep Singh, Shreya Khobragade, Samridhi Hooda"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 5
    collapsed: true
  #theme: lumen
---

```{r load_data_libraries, include=FALSE}
load("saved_data.rda")
library(knitr)
#library(boookdown)
```
### Understanding Emotion in Hemingway's and Fitzgerald's work

### Introduction and Research Question

This project is a computational analysis of emotion in selected twentieth century literary texts. Because of the subjective nature of emotion, it is typically analysed through close reading only. But the use of a data-driven method forces us to think about different questions. What are the ways in which emotions are written? How do they differ across authors, subjects, and times? Which characters, genres and narratives are they most associated with? How does one quantify emotion? Is it possible to quantify it at all, and how does that change the focus of literary studies and stylistic analysis?

We will briefly explore some of these questions in our analysis of F. Scott Fitzgerald and Ernest Hemingway. We aim to use data to find out which author’s work is more emotional, and why? How do these emotions manifest in and around their protagonists?

#### Hypothesis

The novels of Fitzgerald will have more emotion than the novels of Hemingway.

#### Corpus Description

We undertook a comparative study of Ernest Hemingway and F. Scott Fitzgerald’s novels. The corpus was created based on availability. We had to access Hemingway’s works through Libgen, and very few plaintext files were as clean as we wanted them to be. Several files had non-text characters appearing, which would have impacted our analysis. Our initial plan was to get four novels for both authors, since Fitzgerald also has only 4 finished novels. But given the short length of The Old Man and the Sea, we chose to add Garden of Eden to our analysis too, making the total five novels for Hemingway.

Despite this, we believe the comparison will yield interesting results for two reasons. Firstly, both authors wrote in the early 1900s, during what was called the “Jazz Age”, but in very different styles and about very different subject matter. Fitzgerald’s world held soirees, idle wealth, and hedonism; while Hemingway’s held grit, frugality, and violence. Their writing styles are also quite distinct, with Hemingway’s work being characterized by its short, clipped sentences, with few adverbs, and Fitzgerald’s with more verbose, descriptive writing. What they do have in common is a picture of post-war industrialist America, wrought with disillusionment but slowly rebuilding.

In such a dynamic setting, we thought it worthwhile to do a reading of emotion in the works of both authors. Both literature and affect do, after all, lie at the intersection of the individual and the social. Our methodology then becomes two-pronged. One one hand, we conduct a stylistic study of the **writing** of emotions; and on the other, we conduct a contextual analysis of how **characters** in various social worlds feel and interact with these emotions. 
Our texts for each author are as follows:

| Hemingway               | Fitzgerald               |
|-------------------------|--------------------------|
| The Old Man and the Sea | This Side of Paradise    |
| A Farewell to Arms      | The Great Gatsby         |
| For Whom the Bell Tolls | Tender is the Night      |
| The Sun Also Rises      | The Beautiful and Damned |
| The Garden of Eden      |                          |

#### Summary

As mentioned above, we used four texts for Hemingway and five for Fitzgerald to analyse the presence of emotion in their texts. The corpus is close to 800000 words long, which makes it shorter than a safe 1 million cut off, however given our limitations because of our desire to stick to novels by Fitzgerald instead of short stories, our corpus length was limited.

The length of the novels by the authors varied considerably, with two shorter novels by both authors included - The Old Man and the Sea for Hemingway at close to 28000 words, and The Great Gatsby at 48700 words, as well as longer novels.

Fitzgerald's novels have a higher vocabulary density (ratio of words in the novel to unique words in the novel), and sentence length as compared to Hemingway’s. For Fitzgerald's novels, the vocabulary density ranges from 0.097 to 0.122 while for Hemingway's novels this value ranges from 0.049 to 0.099. This indicates the difference in their writing styles and choice of unique words, fitting with our intuition about Fitzgerald's work being more descriptive and using more complex, verbose language. This also applies to the sentence length, with Fitzgerald's average words per sentence ranging from 14.7 to 15.9, while for Hemingway this value ranged from 8.8 to 14.4. This seems reflective of the differences in their writing styles.

With regard to unique terms, in all novels except for The Old Man and the Sea, the unique terms are character names. Interestingly The Old Man and the Sea is the exception to this trend, probably because of the limited human characters in the book, with terms like dolphin and shark standing out as distinctive words.

Thus, our corpus summary seems to indicate that our intuition about the differences between Hemingway’s and Fitzgerald’s writing style seems accurate, and our subsequent sections will focus on how emotions get expressed in these differing writing styles.

### Data Visualization 1: Sentiment Analysis/Relative Frequency

This will determine which author is more emotional. In qualitative reading of literature, it is easy to question what it means for one text to be more emotional than another. One may argue that even simple descriptive paragraphs have emotion attached to them. But it is important to note that our tools will only tell us which author’s work is more overtly emotional: that is, which author’s work will let us decipher emotion without having to analyse subtext or metaphors. Close reading may or may not corroborate our findings, and lack of emotion in our findings definitely does not mean that an author is emotionless.

To answer our question, we created a lexicon with custom words to represent various emotions. These include basic, broad ones like joy and anger, but also more specific ones such as pride and resentment. A function also allows us to pick up on some variations of form in these words. So while our original lexicon has the word “happy”, the function will also include “happiness”, “happier”, “happiest”, etc. by looking for the prefix “happi” and cleaning out all irrelevant results. However, since this lexicon has been manually compiled, there is always the danger of missing out on possible variations.

This gave us the relative percentage of emotion words in each author’s work, as a proportion of all words used in the text.

```{r echo=FALSE}
data_vis_1_int
```

As the graph indicates, Fitzgerald scores much higher on the occurrences of emotion words. His least emotional novel, The Great Gatsby, with 2.5%, is still higher than Hemingway's most emotional novel, A Farewell To Arms, which is around 2.1%. The Old Man and The Sea has the lowest score of the corpus.

The reason for Hemingway's scores is largely stylistic. Because a lot of emotion is directly located in adverbs, a good percentage of our lexicon consists of adverbs, which are exactly what Hemingway's writing lacks. A lot of Hemingway's characters were also hypermasculine and shied away from being upfront about their feelings. In fact, if one were to read closely, it is by means of this stoic masculine facade that their true feelings can be deciphered. Our tools are not equipped to do this. They look at words rather than sentences or paragraphs, so they lack any reading whatsoever of subtext or context. They cannot pick up on analogies, or metaphors, or the manner in which authors choose to express characters' emotions through their expressions, or reactions.

Fitzgerald's relatively higher percentage is reflective of his writing style, and the larger themes in his novels. His writing style is more descriptive, and allows for more space to explore various emotions. His novels also focus on the emotional bonds between characters, especially within the context of romantic relationships, which is a likely reason for the higher score too. Several of his male main characters have romantic relationships which provide a space for the important themes in the book to emerge through, making it important for the writing and descriptions of the same to be detailed.

### Data Visualization 2: Specific Emotions in Authors

For our next step, we used the NRC emotion lexicon to determine which emotions were highest. This would ideally give us an insight into the characters’ social worlds, as well as the thematic foci of the authors.

To avoid an overcrowded graph, we picked four emotions from the positive end of the spectrum (trust, anticipation, surprise, and joy -- although surprise can go both ways) and four from the negative end (anger, sadness, fear, and disgust). Unlike our previous lexicon, this lexicon was pre-made and external, and it could have missed out on certain terms.

```{r echo=FALSE}
data_vis_2
```

Here, Hemingway scores significantly higher on anger and fear, and marginally higher on disgust, trust and sadness. All four of the “negative” emotions are higher in his work. Intuitively, this makes sense for anger and fear, because he writes characters with lots of machismo and puts them in demanding settings like war or at the mercy of nature. _A Farewell To Arms_ features a protagonist who suffers the pain of war and is disillusioned after it; while The Sun Also Rises speaks of how the protagonists’ experiences at war cripple his romances as well. We can justify his usage of trust too, because people in these settings often develop trust just by virtue of surviving their hardships together (like in _The Old Man and The Sea_).

What’s interesting (and not intuitive) is that Fitzgerald has barely any words that express disgust. One would assume that it is such an easy emotion for snobbish upper class people to feel -- especially when they are continually in unhappy love affairs. The settings within which there can be feelings of disgust emerging are not less- there are characters who undergo a “fall”, and resort to alcoholism, debauchery, and self-pity due to failures in relationships, or characters who cheat and betray their significant others. Despite this, there is little to no disgust, even though there are several situations which would provide ample space for the expression of disgust. Perhaps these results are due to us conflating disgust and shame, or another related emotion that may show up in future tests.

### Data Visualization 3: Named Entity Recognition

While NRC gives us a powerful picture of which emotions dominate an author’s work, we wanted to further inquire into where and who they come from. Hence, we combined this with Named Entity Recognition (NER) and sentiment analysis to map the emotions surrounding the characters in each novel. For the protagonists, we retained the emotion lexicon, so we could pinpoint specific emotions; and for the other characters, we simply generated a positive or negative sentiment score.

Who the protagonist is in a story can be unclear, but we’ve used the safest assumption based on research. This analysis would be more difficult with a larger corpus because of how time consuming the cleaning process for NER is. The amount we learn about each protagonist varies, with some novels providing great insights into their emotions, and some such as _The Great Gatsby_, emphasising on the same less.

Hemingway’s protagonists seem to consistently have higher scores for trust, as compared to other emotions. Interestingly, despite his protagonists being presented as extremely “masucline” characters, thrust in situations of wars, or grappling with nature, they score higher on trust than they do on anger. No characters are particularly angry: instead, sadness and trust, and in one case fear too, stand out more. This brings up an interesting point about these characters and the manner in which Hemingway thought about masculinity in the post war era he was writing in. Trust is an emotion that is relational - it emerges within the context of relationships, or a cause, and is thus an indicator of the same in novels. As demonstrated by characters such as Henry from _A Farewell to Arms_, the emphasis on trust could be a product of the relationships, and the emphasis on men `finding themselves’ and in the case of the novel, bidding goodbye to war to embrace a life of love instead. Thus, the comparatively high scores for trust could be a product of the relationships exhibited in Hemingway’s novels. Hemingway’s choice to make his main characters masculine, but also not intensely angry allows him to create characters who experience complex emotional conflict, and thus, cannot be typecast as one dimensional characters.

However, despite this, Hemingway's characters are angrier than Fitzgerald’s (except for Dick Diver), which fits with our assumptions about them. It is possible that Hemingway may not have had very angry characters in absolute terms, but as compared to other authors of his time, he may have written people that are angrier than the average character. This is not accounted for in our analysis.

Fitzgerald’s protagonists score high on trust and anticipation as compared to other emotions. This is likely because of the emphasis on romantic relationships in his books, and the way they present these relationships as causing the downfall of the male protagonist - emphasising on the presence (or absence) of trust in these relationships. The high score for anticipation could be a product of two factors - first, the male character’s yearning for a particular female character, and second, the build up to the protagonist’s downfall. The main characters longing for a female character can be seen in the _The Great Gatsby_, where Gatsby would hold parties in anticipation of Daisy coming to them. Similarly, the build up to the characters’ downfall can be visible in _Tender is the Night_, where the reader is witness to Dick Diver’s slow fall from grace.

His protagonists also feel more joy than sadness (barring Dick Diver), and this stands out given our initial assumption that Fitzegerald’s novels would have more melancholic characters. However, apart from Gatsby, none of them particularly feel surprise, which is, if we may, quite surprising. Partly because the protagonist of any story typically has unforeseen events happen to them, and thus is surprised to at least some extent; and partly because a lot of Fitzgerald’s protagonists go through similar situations as Gatsby, in terms of being abandoned by a lover. Perhaps it is Gatsby’s murder that skews the results, or perhaps the other protagonists are simply not surprised by these events, and instead anticipated them, given the reality of their lives.

Our analysis also includes a list of the most positive and negative characters in each novel, created through sentiment analysis that generated a score for each of them. Scores on the positive end mean the character had positive emotions associated with them, while negative scores mean the opposite. This provides insights into how emotion and gender was linked. Fitzgerald has more negative women than positive women. This could be reflective of how his women influenced the protagonists’ lives, such as in the case of _The Great Gatsby_ and _Tender is the Night_, where they’re presented as leading to the protagonist's downfall. He presents these romantic relationships as being all consuming, with the woman either cheating on the protagonist, or even indirectly causing his death. Some of Fitzgerald’s female characters, such as Nicole Diver are also presented with a history of mental illness, contributing to the assignment of negative sentiments to them.

Hemingway on the other hand, has more positive women. This could be reflective of the idealised image of women his protagonists have, where they are put on moral pedestals that also become cages. Some women of Hemingway were incredibly submissive and obedient, which he saw as a thing of great value. One such character is Catherine from _A Farewell to Arms_ -- but interestingly, she does not appear in the list of most positive characters from the novel. The two women that do appear are her colleagues, who are secondary to the main romance. This may have been due to the lexicon classifying a lot of what Catherine did as neutral, rather than positive or negative. It is always an error that can exist with characters and dialogue in sentiment analysis.

### Conclusion

The research proves our hypothesis, and goes on to further complicate it. Our hypothesis was proven correct, as Fitzgerald does have more emotion than Hemingway, as data has shown. He uses a higher percentage of emotion words, and as we assumed, this seems to be a product of their respective writing styles. However, there are several caveats to the same as we use a very narrow view of emotion with our emotion lexicons.

As our analysis reveals, Hemingway is not emotionless by any means. Instead, his writing style is the primary reason because of which he appears to express less emotion in his writing. Thus, our hypothesis assessed emotion through a narrow lens and a better way of understanding it would be that Fitzgerald wrote with more easily identifiable emotion words than Hemingway, rather than assuming Hemingway’s characters lacked emotional depth. . They also helped to counter some initial fleeting assumptions we made about the corpus, like Fitzgerald’s characters being sadder or Hemingway’s characters being one-dimensionally angry, and allowed for a more nuanced understanding of their characters.

### Reflection

R provided us with tools to view the texts and an author's work at a macro level, and allowed us to deal with a large quantity of data in a short period of time. This was useful as it presented trends across novels for different authors. The visualisations also allowed for the easy analysis and presentation of findings. Given that the main rule of writing is to show, not tell readers how a character feels, one may wonder if these are worthwhile tools to explore emotion at all, considering they only pick up on what is explicitly told to them. But they are very powerful tools to analyse writing style as well as emotion in relation to other variables in the text, like gender or character.

The activity involved moving away from a typical close reading of literature, which presented its own challenges. A significant issue revolved around how emotion lexicons and sentiment scores are not useful for picking up on subtext. It cannot measure emotions expressed through metaphors, details of non verbal expressions in books and some dialogues. It is more useful at identifying more overt methods of expressing emotion. Along with this, it is difficult for even the most elaborate emotional lexicons to be entirely comprehensive, indicating the possibility that several expressions of emotion could have been missed.

Another issue with our analysis was related to how we generated an overall emotion score, instead of understanding sentiment around a concept (such as sentiment around war in novels). This proved to be slightly broad, as compared to drilling down on a topic, however, it was appropriate for our hypothesis which sought to explore emotion in their writing styles.

### 2021-10-14 project_2.Rmd
Final HTML

@@
 library(scales)
+library(plotly)
+library(htmlwidgets)
 ```
@@
 
-ggRadar(
+data_vis_2 <- ggRadar(
   data = sentiment_both,
@@
 
-p<-ggplot(data=combined_table, aes(x=percent, y=title, fill=author)) +
-  geom_bar(stat="identity")
-p
+data_vis_1 <-ggplot(data=combined_table, aes(x=percent, y=title, fill=author, text = paste("% of emotion: ", percent, sep = ""))) +
+  geom_bar(stat="identity") + labs(x = "% of emotion detected", 
+       y = "All novels", fill = "Author", title = "Emotive content in all novels")
+
+data_vis_1_int <- ggplotly(data_vis_1, tooltip = "text")
+data_vis_1_int
 ```
@@
 ```{r exporting_tables}
-save(author_table, hemingway_corpus, fitzgerald_corpus, file = "saved_data.rda")
+save(author_table, hemingway_corpus, fitzgerald_corpus, data_vis_1_int, data_vis_2, file = "saved_data.rda")
 ```

### 2021-10-14 NER_analysis_final.Rmd
Error correction
Added final HTML. Project complete.

@@
 full_df_backup <- full_df
+write_csv(full_df, "full_df_backup.csv")
 ```
@@
 ```{r message=FALSE}
-ner_total_sentiment %>%
+vis_ner_1 <- ner_total_sentiment %>%
   group_by(title) %>%
-  filter(kind == "person") %>%
+  filter(kind == "person") %>% filter(title != "The Old Man and the Sea") %>%
   top_n(5) %>%
   mutate(words = reorder(words, total)) %>%
-  ggplot(aes(words, y = total, fill = title)) +
+  ggplot(aes(words, y = total, fill = title, text = paste("Character: ", words, "\n", "% of positive emotion: ", total, sep = ""))) +
   geom_col()  +
-  facet_wrap(~ title, scales = "free") +
-  coord_flip()
+  facet_wrap(~ title, scales = "free", ncol = 2) +coord_flip() + labs(x = "Names of characters", 
+       y = "% of positive emotion", fill = "Novel", title = "% of Positive Emotions Detected")
+
+vis_ner_1_int <- ggplotly(vis_ner_1, tooltip = "text")
+vis_ner_1_int 
 ```
-
+theme(plot.margin = margin(0, 0, 0, 0, "cm"))
 ### Chart of the top 10 negative entities
@@
 ```{r message=FALSE}
-ner_total_sentiment %>%
+vis_ner_2 <- ner_total_sentiment %>%
   group_by(title) %>%
@@
   mutate(words = reorder(words, (desc(total)))) %>%
-  ggplot(aes(words, y = total, fill = title)) +
+  ggplot(aes(words, x = total, fill = title, text = paste("Character: ", words, "\n", "% of negative emotion: ", 0 - total, sep = ""))) +
   geom_col() +
-  facet_wrap(~ title, scales = "free") +
-  coord_flip()
+  facet_wrap(~ title, scales = "free") + labs(x = "Names of characters", 
+       y = "% of negative emotion", fill = "Novel", title = "% of Negative Emotions Detected")
+vis_ner_2_int <- ggplotly(vis_ner_2, tooltip = "text")
+vis_ner_2_int
+
 ```
@@
 
-```{r message=FALSE graphs_fitzgerald}
+```{r message=FALSE}
 radar_facet_sentiment <- entities_matches_sentiment %>%
@@
 
-sentiment_Amory <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Amor"), ] %>% filter(title == "This Side of Paradise") %>%
-  group_by(title, words, kind) %>%
-  summarise(across(anger:sentiment, sum)) %>% select(-positive, -negative, -sentiment, -kind)
-
-sentiment_Henry <- entities_matches_sentiment[str_detect(entities_matches_sentiment$words, "Henry"), ] %>% filter(title == "A Farewell to Arms") %>%
+sentiment_Amory <- entities_matches_sentiment %>% filter(words == "Amory") %>% filter(title == "This Side of Paradise") %>%
   group_by(title, words, kind) %>%
@@
 
-ggRadar(
+vis_ner_3 <- ggRadar(
 
@@
 )
+vis_ner_3
 ```
@@
 
-ggRadar(
+vis_ner_4 <- ggRadar(
 
@@
 )
+vis_ner_4
 
 ```
+
+```{r exporting_tables}
+save(vis_ner_1_int, vis_ner_2_int, vis_ner_3, vis_ner_4, file = "saved_ner.rda")
+```

### 2021-10-14 Submission.Rmd
Added final HTML. Project complete.

@@
 load("saved_data.rda")
+load("saved_ner.rda")
 library(knitr)
-#library(boookdown)
 ```
@@
 This project is a computational analysis of emotion in selected twentieth century literary texts.
-Because of the subjective nature of emotion, it is typically analysed through close reading only.
+Owing to the subjective nature of emotion, it is typically analysed through close reading of texts only.
 But the use of a data-driven method forces us to think about different questions.
-What are the ways in which emotions are written?
-How do they differ across authors, subjects, and times?
+What are the ways in which emotions are _written_?
+How do they _differ_ across authors, subjects, and times?
 Which characters, genres and narratives are they most associated with?
-How does one quantify emotion?
-Is it possible to quantify it at all, and how does that change the focus of literary studies and stylistic analysis?
+How does one _quantify_ emotion?
+If emotion is indeed quantifiable, how does that change the focus of literary studies and stylistic analysis?
 
@@
 Scott Fitzgerald and Ernest Hemingway.
-We aim to use data to find out which author’s work is more emotional, and why?
-How do these emotions manifest in and around their protagonists?
+We aim to use data to find out which author’s work is more emotional, and why.
+We are also interested in exploring how these emotions manifest in and around the protagonists of these novels.
 
@@
 
-The novels of Fitzgerald will have more emotion than the novels of Hemingway.
+The novels of F. Scott Fitzgerald contain more emotion than the novels of Earnest Hemingway.
 
@@
 The corpus was created based on availability.
-We had to access Hemingway’s works through Libgen, and very few plaintext files were as clean as we wanted them to be.
-Several files had non-text characters appearing, which would have impacted our analysis.
-Our initial plan was to get four novels for both authors, since Fitzgerald also has only 4 finished novels.
-But given the short length of The Old Man and the Sea, we chose to add Garden of Eden to our analysis too, making the total five novels for Hemingway.
+Hemingway's works were not on available on Gutenberg, and had to be obtained from a different resource.
+The files thus obtained were not clean, and contained textual artifacts.
+The files were cleaned thoroughly so as to prevent issues in analysis.
+Since Fitzgerald has only 4 finished novels, we obtained 4 novels of Hemingway as well.
+But the word count differed drastically, so we added Hemingway's _Garden of Eden_ to reach desirable balance in wordcount.
 
-Despite this, we believe the comparison will yield interesting results for two reasons.
-Firstly, both authors wrote in the early 1900s, during what was called the “Jazz Age”, but in very different styles and about very different subject matter.
+Despite this minor setback, we were optimistic that we would find interesting results, for two reasons.
+Firstly, both authors wrote in the early 1900s, during what was called the “Jazz Age”, but in very different styles, and about very different subject matter.
 Fitzgerald’s world held soirees, idle wealth, and hedonism; while Hemingway’s held grit, frugality, and violence.
-Their writing styles are also quite distinct, with Hemingway’s work being characterized by its short, clipped sentences, with few adverbs, and Fitzgerald’s with more verbose, descriptive writing.
+Their writing styles are also quite distinct, with Hemingway’s work being characterized by its short, clipped sentences, devoid of adverbs, and Fitzgerald’s with more verbose, descriptive writing.
 What they do have in common is a picture of post-war industrialist America, wrought with disillusionment but slowly rebuilding.
@@
 Both literature and affect do, after all, lie at the intersection of the individual and the social.
-Our methodology then becomes two-pronged.
-One one hand, we conduct a stylistic study of the **writing** of emotions; and on the other, we conduct a contextual analysis of how **characters** in various social worlds feel and interact with these emotions.
+Our methodology then became two-pronged.
+One one hand, we conducted a stylistic study of the **writing** of emotions; and on the other, we conducted a contextual analysis of how **characters** in various social worlds feel and interact with these emotions.
 Our texts for each author are as follows:
@@
 
-As mentioned above, we used four texts for Hemingway and five for Fitzgerald to analyse the presence of emotion in their texts.
-The corpus is close to 800000 words long, which makes it shorter than a safe 1 million cut off, however given our limitations because of our desire to stick to novels by Fitzgerald instead of short stories, our corpus length was limited.
+We used five texts for Hemingway and four for Fitzgerald to analyse the presence of emotion in their texts.
+The corpus is close to _800,000_ words long, which is just shy of the 1 million mark considered adequate.
+Our choice of the authors limited our options, but we excitedly proceeded with our corpus.
 
-The length of the novels by the authors varied considerably, with two shorter novels by both authors included - The Old Man and the Sea for Hemingway at close to 28000 words, and The Great Gatsby at 48700 words, as well as longer novels.
+The length of the novels by the authors varied considerably, with two shorter novels by both authors included - _The Old Man and the Sea_ for Hemingway, at approximately _28,000_ words, and _The Great Gatsby_ at _48,700_ words, as well as longer novels.
 
 Fitzgerald's novels have a higher vocabulary density (ratio of words in the novel to unique words in the novel), and sentence length as compared to Hemingway’s.
-For Fitzgerald's novels, the vocabulary density ranges from 0.097 to 0.122 while for Hemingway's novels this value ranges from 0.049 to 0.099.
-This indicates the difference in their writing styles and choice of unique words, fitting with our intuition about Fitzgerald's work being more descriptive and using more complex, verbose language.
-This also applies to the sentence length, with Fitzgerald's average words per sentence ranging from 14.7 to 15.9, while for Hemingway this value ranged from 8.8 to 14.4.
+For Fitzgerald's novels, the vocabulary density ranges from _0.097_ to _0.122_ while for Hemingway's novels this value ranges from _0.049_ to _0.099_.
+This indicates that Hemingway used a more comfortable vocabulary, while Fitzgerald incorporated more unique words in his writing.
+The difference in their writing styles and choice of unique words fit well with our intuition about Fitzgerald's work being more descriptive and employing more complex, verbose language.
+This also applies to the sentence length, with Fitzgerald's average words per sentence ranging from _14.7_ to _15.9_, while for Hemingway this value ranged from _8.8_ to _14.4_.
 This seems reflective of the differences in their writing styles.
 
-With regard to unique terms, in all novels except for The Old Man and the Sea, the unique terms are character names.
-Interestingly The Old Man and the Sea is the exception to this trend, probably because of the limited human characters in the book, with terms like dolphin and shark standing out as distinctive words.
+With regard to unique terms, in all novels except for _The Old Man and the Sea_, the most unique terms are character names.
+Interestingly, _The Old Man and the Sea_ is the exception to this trend, probably because of the limited human characters in the book, with terms like dolphin and shark standing out as distinctive words.
 
-Thus, our corpus summary seems to indicate that our intuition about the differences between Hemingway’s and Fitzgerald’s writing style seems accurate, and our subsequent sections will focus on how emotions get expressed in these differing writing styles.
+Thus, our corpus summary seems to indicate that our intuition about the differences between Hemingway’s and Fitzgerald’s writing style seems accurate, and our subsequent sections will focus on how emotions are expressed in these differing writing styles.
 
-### Data Visualization 1: Sentiment Analysis/Relative Frequency
+### Data Visualization 1: Sentiment Analysis and Relative Frequency
 
-This will determine which author is more emotional.
 In qualitative reading of literature, it is easy to question what it means for one text to be more emotional than another.
 One may argue that even simple descriptive paragraphs have emotion attached to them.
-But it is important to note that our tools will only tell us which author’s work is more overtly emotional: that is, which author’s work will let us decipher emotion without having to analyse subtext or metaphors.
-Close reading may or may not corroborate our findings, and lack of emotion in our findings definitely does not mean that an author is emotionless.
+But it is important to note that our tools are only equipped tell us which author’s work is more overtly emotional: that is, which author’s work will let us decipher emotion without having to analyse subtext or metaphors.
+Close reading may or may not corroborate our findings, and lack of emotion in our findings definitely does not mean that an author employs less emotion in his writing.
 
-To answer our question, we created a lexicon with custom words to represent various emotions.
+For this task, we created a lexicon with handpicked words to represent various emotions.
 These include basic, broad ones like joy and anger, but also more specific ones such as pride and resentment.
-A function also allows us to pick up on some variations of form in these words.
-So while our original lexicon has the word “happy”, the function will also include “happiness”, “happier”, “happiest”, etc.
-by looking for the prefix “happi” and cleaning out all irrelevant results.
-However, since this lexicon has been manually compiled, there is always the danger of missing out on possible variations.
+We also employed code to generate variations of these words, including adverbs, and thus we had a sizable number of unique emotion words at our disposal.
+However, since this lexicon has been manually compiled, the possibility that some useful words were missed always lingers.
 
-This gave us the relative percentage of emotion words in each author’s work, as a proportion of all words used in the text.
+We first matched these words with each author's corpus, and calculated their occurence.
+To get a more accurate picture, we further divided this number by the total number of words in each author's corpus.
+The relative frequency provides interesting insights.
 
@@
 
-As the graph indicates, Fitzgerald scores much higher on the occurrences of emotion words.
-His least emotional novel, The Great Gatsby, with 2.5%, is still higher than Hemingway's most emotional novel, A Farewell To Arms, which is around 2.1%.
-The Old Man and The Sea has the lowest score of the corpus.
+As the graph indicates, Fitzgerald work has a higher percentage of these emotion words.
+His least emotional novel, _The Great Gatsby_, with only _2.5%_ of the words matching our lexicon, is still higher than Hemingway's most emotional novel -- _A Farewell To Arms_, which stands at _2.1%_.
+_The Old Man and The Sea_ has the lowest score of the corpus.
 
 The reason for Hemingway's scores is largely stylistic.
-Because a lot of emotion is directly located in adverbs, a good percentage of our lexicon consists of adverbs, which are exactly what Hemingway's writing lacks.
-A lot of Hemingway's characters were also hypermasculine and shied away from being upfront about their feelings.
+Because a lot of emotion is embedded in adverbs, a good percentage of our lexicon consists of adverbs, which are exactly what Hemingway's writing lacks.
+A lot of Hemingway's characters are also hypermasculine and shy away from explicit expression of their emotions.
 In fact, if one were to read closely, it is by means of this stoic masculine facade that their true feelings can be deciphered.
-Our tools are not equipped to do this.
-They look at words rather than sentences or paragraphs, so they lack any reading whatsoever of subtext or context.
-They cannot pick up on analogies, or metaphors, or the manner in which authors choose to express characters' emotions through their expressions, or reactions.
+Our tools, however, lack the ability to detect subtext and decipher metaphors, analogies, irony, sarcasm, and the likes.
+Even when reading at a sentence- or paragraph-level, the tools are ill-equipped to decode a non-explicit expression of emotion.
 
 Fitzgerald's relatively higher percentage is reflective of his writing style, and the larger themes in his novels.
-His writing style is more descriptive, and allows for more space to explore various emotions.
-His novels also focus on the emotional bonds between characters, especially within the context of romantic relationships, which is a likely reason for the higher score too.
+His writing style is more descriptive, and he freely employs complex expressions that stitch multiple metaphors together.
+His novels also focus on the emotional bonds between characters, especially within the context of romantic relationships, which may have contributed to his higher score.
 Several of his male main characters have romantic relationships which provide a space for the important themes in the book to emerge through, making it important for the writing and descriptions of the same to be detailed.
@@
 
-For our next step, we used the NRC emotion lexicon to determine which emotions were highest.
+Now that we had obtained evidence in favor of our hypothesis, we used the NRC emotion lexicon to determine which emotions had the highest expression in the texts.
 This would ideally give us an insight into the characters’ social worlds, as well as the thematic foci of the authors.
 
-To avoid an overcrowded graph, we picked four emotions from the positive end of the spectrum (trust, anticipation, surprise, and joy -- although surprise can go both ways) and four from the negative end (anger, sadness, fear, and disgust).
-Unlike our previous lexicon, this lexicon was pre-made and external, and it could have missed out on certain terms.
+To avoid an overcrowded graph, we picked four emotions from the positive end of the spectrum ( _trust_, _anticipation_, _surprise_, and _joy_ -- although _surprise_ can belong to either category) and four from the negative end ( _anger_, _sadness_, _fear_, and _disgust_).
+Unlike our previous lexicon, this lexicon was pre-made and external, and had its own limitations.
+Each emotion had attached to itself a list of words related to the respective emotion, with some words overlapping among the emotions.
+We calculated the expression of each emotion in the corpus of each author. Then we relativised each emotion so we could compare them fairly.
 
@@
 ```
-
-Here, Hemingway scores significantly higher on anger and fear, and marginally higher on disgust, trust and sadness.
-All four of the “negative” emotions are higher in his work.
+Here, Hemingway scores significantly higher on _anger_ and _fear_, and marginally higher on _disgust_, _trust_ and _sadness_.
+All four of the “negative” emotions find higher expression in his work.
 Intuitively, this makes sense for anger and fear, because he writes characters with lots of machismo and puts them in demanding settings like war or at the mercy of nature.
-_A Farewell To Arms_ features a protagonist who suffers the pain of war and is disillusioned after it; while The Sun Also Rises speaks of how the protagonists’ experiences at war cripple his romances as well.
+_A Farewell To Arms_ features a protagonist who suffers the pain of war and is disillusioned after it; while _The Sun Also Rises_ speaks of how the protagonist's experiences at war cripple his romances as well.
 We can justify his usage of trust too, because people in these settings often develop trust just by virtue of surviving their hardships together (like in _The Old Man and The Sea_).
 
-What’s interesting (and not intuitive) is that Fitzgerald has barely any words that express disgust.
+What’s interesting (and not intuitive) is that Fitzgerald uses surprisingly few words that express _disgust_.
 One would assume that it is such an easy emotion for snobbish upper class people to feel -- especially when they are continually in unhappy love affairs.
-The settings within which there can be feelings of disgust emerging are not less- there are characters who undergo a “fall”, and resort to alcoholism, debauchery, and self-pity due to failures in relationships, or characters who cheat and betray their significant others.
-Despite this, there is little to no disgust, even though there are several situations which would provide ample space for the expression of disgust.
-Perhaps these results are due to us conflating disgust and shame, or another related emotion that may show up in future tests.
+The settings within which feelings of disgust can emerge are not few -- there are characters who undergo a “fall”, and resort to alcoholism, debauchery, and self-pity due to failures in relationships, or characters who cheat and betray their significant others.
+Despite this, there is little to no expression of _disgust_, even though there are several situations which would provide ample space for the expression of it.
+Perhaps these results are due to us conflating _disgust_ and _shame_, or another related emotion that may show up in future tests.
+It is also important to note here that the NRC lexicon contains no adverbs, and since the computation involved only exact matches, the analysis was deprived of adverbs which, if accounted for, may have produced different results.
 
@@
 
-While NRC gives us a powerful picture of which emotions dominate an author’s work, we wanted to further inquire into where and who they come from.
-Hence, we combined this with Named Entity Recognition (NER) and sentiment analysis to map the emotions surrounding the characters in each novel.
-For the protagonists, we retained the emotion lexicon, so we could pinpoint specific emotions; and for the other characters, we simply generated a positive or negative sentiment score.
+While sentiment analysis using NRC provided us a powerful picture of which emotions dominate an author’s work, we wanted to further inquire into where and who they originated from.
+Hence, we performed Named Entity Recognition to locate the persons in the text.
+Then we performed sentiment analysis on the text surrounding these entities to find emotional valences.
+We employed the NRC lexicon again, because we wanted to maintain consistency with prior analyses, and NRC was best suited for our needs because of its comprehensive lexicon.
 
-Who the protagonist is in a story can be unclear, but we’ve used the safest assumption based on research.
-This analysis would be more difficult with a larger corpus because of how time consuming the cleaning process for NER is.
-The amount we learn about each protagonist varies, with some novels providing great insights into their emotions, and some such as _The Great Gatsby_, emphasising on the same less.
+It is difficult to extract the protagonist through computational means, so we relied on external resources to compile a list of protagonists of each novel.
+This method is not suited for a corpus that has thousands of books, but since we had only 9 works in our corpus, we chose this route.
+The NER tagging worked well for extracting entities, but the text had to be manually scrubbed line-by-line to fix the errors.
+The amount we learned about each protagonist varied, with some novels providing great insights into their emotional content, but some, such as _The Old Man and The Sea_ still withheld their secrets.
+```{r echo=FALSE}
+vis_ner_4
+```
+Hemingway’s protagonists appear to consistently have higher scores for _trust_, as compared to other emotions.
+Interestingly, despite his protagonists being presented as extremely “masculine” characters, thrust in situations of wars, or grappling with the dangers of nature, they score higher on _trust_ than they do on _anger_.
+On the whole, the characters seem well balanced, with one exception -- _Henry_.
+_Sadness_ and _trust_, and in one case _fear_ appear to be prominent emotions.
+This yields an interesting insight about these characters and the manner in which Hemingway thought about masculinity in the post war era he was writing in.
+_Trust_ is an emotion that is relational - it emerges within the context of relationships, or a cause, and is thus an indicator of the same in novels.
+As demonstrated by characters such as _Henry_ from _A Farewell to Arms_, the emphasis on trust could be a product of the relationships, and the emphasis on men 'finding themselves’ and in the case of the novel, bidding goodbye to war to embrace a life of love instead.
+Thus, the comparatively high scores for trust could be a product of the relationships exhibited in Hemingway’s novels.
+Hemingway’s choice to make his main characters masculine, but also not intensely angry, allows him to create characters who experience complex emotional conflict, and thus, cannot be typecast as one-dimensional characters.
 
-Hemingway’s protagonists seem to consistently have higher scores for trust, as compared to other emotions.
-Interestingly, despite his protagonists being presented as extremely “masucline” characters, thrust in situations of wars, or grappling with nature, they score higher on trust than they do on anger.
-No characters are particularly angry: instead, sadness and trust, and in one case fear too, stand out more.
-This brings up an interesting point about these characters and the manner in which Hemingway thought about masculinity in the post war era he was writing in.
-Trust is an emotion that is relational - it emerges within the context of relationships, or a cause, and is thus an indicator of the same in novels.
-As demonstrated by characters such as Henry from _A Farewell to Arms_, the emphasis on trust could be a product of the relationships, and the emphasis on men `finding themselves’ and in the case of the novel, bidding goodbye to war to embrace a life of love instead.
-Thus, the comparatively high scores for trust could be a product of the relationships exhibited in Hemingway’s novels.
-Hemingway’s choice to make his main characters masculine, but also not intensely angry allows him to create characters who experience complex emotional conflict, and thus, cannot be typecast as one dimensional characters.
+However, despite this, Hemingway's characters express more anger than Fitzgerald’s (except for Dick Diver), which fits with our assumptions about them.
+It is possible that Hemingway's characters may not have a higher expression of anger when compared to his peers', but for our narrow hypothesis they do appear to be more highly associated with _anger_ than Fitzgerald's characters.
+```{r echo=FALSE}
+vis_ner_3
+```
+Fitzgerald’s protagonists score high on _trust_ and _anticipation_ as compared to other emotions.
+This is likely the result of a focus on romantic relationships in his books.
+Further, these relationships are presented as causing the downfall of the male protagonist -- emphasising on the presence (or absence) of trust in these relationships.
+The high score for anticipation could be a product of two factors -- first, the male character’s yearning for a particular female character, and second, the build up to the protagonist’s downfall.
+The main characters' longing for a female character can be seen in the _The Great Gatsby_, where Gatsby would hold parties in anticipation of Daisy coming to them.
+Similarly, the build up to the characters’ downfall can be observed in _Tender is the Night_, where the reader is witness to Dick Diver’s slow fall from grace.
 
-However, despite this, Hemingway's characters are angrier than Fitzgerald’s (except for Dick Diver), which fits with our assumptions about them.
-It is possible that Hemingway may not have had very angry characters in absolute terms, but as compared to other authors of his time, he may have written people that are angrier than the average character.
-This is not accounted for in our analysis.
-
-Fitzgerald’s protagonists score high on trust and anticipation as compared to other emotions.
-This is likely because of the emphasis on romantic relationships in his books, and the way they present these relationships as causing the downfall of the male protagonist - emphasising on the presence (or absence) of trust in these relationships.
-The high score for anticipation could be a product of two factors - first, the male character’s yearning for a particular female character, and second, the build up to the protagonist’s downfall.
-The main characters longing for a female character can be seen in the _The Great Gatsby_, where Gatsby would hold parties in anticipation of Daisy coming to them.
-Similarly, the build up to the characters’ downfall can be visible in _Tender is the Night_, where the reader is witness to Dick Diver’s slow fall from grace.
-
-His protagonists also feel more joy than sadness (barring Dick Diver), and this stands out given our initial assumption that Fitzegerald’s novels would have more melancholic characters.
-However, apart from Gatsby, none of them particularly feel surprise, which is, if we may, quite surprising.
-Partly because the protagonist of any story typically has unforeseen events happen to them, and thus is surprised to at least some extent; and partly because a lot of Fitzgerald’s protagonists go through similar situations as Gatsby, in terms of being abandoned by a lover.
-Perhaps it is Gatsby’s murder that skews the results, or perhaps the other protagonists are simply not surprised by these events, and instead anticipated them, given the reality of their lives.
-
-Our analysis also includes a list of the most positive and negative characters in each novel, created through sentiment analysis that generated a score for each of them.
+His protagonists also feel more _joy_ than _sadness_ (barring Dick Diver), and this contrasts our assumption that Fitzegerald’s novels would have more melancholic characters.
+However, apart from Gatsby, none of them particularly feel _surprise_, which is, if we may, quite surprising.
+Partly because the protagonist of any story typically has unforeseen events happen to them, and thus is surprised to at least some extent; and partly because many of Fitzgerald’s protagonists go through similar situations as Gatsby, in terms of being abandoned by a lover.
+Perhaps it is Gatsby’s murder that skews the results, or perhaps the other protagonists are simply not surprised by these events.
+It is also likely that _surprise_ may not be an emotion that has strong associations with the words used to capture it.
+For example, the phrase "Oh my God", stripped of "Oh", may not be accounted for, especially when the sentiments focus on unigrams.
+A better approach would be to include bigrams, but that is also true for other emotions.
+```{r echo=FALSE}
+vis_ner_1_int
+```
+Our analysis also produced a list of the most positive and negative characters in each novel, created through sentiment analysis that generated a score for each of them.
 Scores on the positive end mean the character had positive emotions associated with them, while negative scores mean the opposite.
-This provides insights into how emotion and gender was linked.
-Fitzgerald has more negative women than positive women.
-This could be reflective of how his women influenced the protagonists’ lives, such as in the case of _The Great Gatsby_ and _Tender is the Night_, where they’re presented as leading to the protagonist's downfall.
+This provides insights into how **emotion** and **gender** was linked.
+Fitzgerald has more _negative_ women than _positive_ women.
+This could be reflective of how his female characters influenced the protagonists’ lives, such as in the case of _The Great Gatsby_ and _Tender is the Night_, where they’re presented as the cause for the protagonist's downfall.
 He presents these romantic relationships as being all consuming, with the woman either cheating on the protagonist, or even indirectly causing his death.
-Some of Fitzgerald’s female characters, such as Nicole Diver are also presented with a history of mental illness, contributing to the assignment of negative sentiments to them.
-
-Hemingway on the other hand, has more positive women.
+Some of Fitzgerald’s female characters, such as _Nicole Diver_ are also presented with a history of mental illness, which contributes to the assignment of negative sentiments to them.
+```{r echo=FALSE}
+vis_ner_2_int
+```
+Hemingway, on the other hand, has more positive female characters.
 This could be reflective of the idealised image of women his protagonists have, where they are put on moral pedestals that also become cages.
-Some women of Hemingway were incredibly submissive and obedient, which he saw as a thing of great value.
-One such character is Catherine from _A Farewell to Arms_ -- but interestingly, she does not appear in the list of most positive characters from the novel.
+Some female characters in Hemingway were incredibly submissive and obedient, which he saw as a thing of great value.
+One such character is _Catherine_ from _A Farewell to Arms_ -- but interestingly, she does not appear in the list of most positive characters from the novel.
 The two women that do appear are her colleagues, who are secondary to the main romance.
-This may have been due to the lexicon classifying a lot of what Catherine did as neutral, rather than positive or negative.
-It is always an error that can exist with characters and dialogue in sentiment analysis.
+It may be worth exploring this further, through computing correlations, to see why secondary characters have more emotion surrounding them.
+How sentences are broken down and associated with entities plays a very important role in this technique of analysis.
 
@@
 
-The research proves our hypothesis, and goes on to further complicate it.
-Our hypothesis was proven correct, as Fitzgerald does have more emotion than Hemingway, as data has shown.
-He uses a higher percentage of emotion words, and as we assumed, this seems to be a product of their respective writing styles.
-However, there are several caveats to the same as we use a very narrow view of emotion with our emotion lexicons.
+Our research has shown that Fitzgerald's works have more expression of emotion than Hemingway's.
+The hypothesis has held its ground despite all the different ways of analysis that were hurled at it.
+The first result, generated by using a curated list of words, shows that, assuming the lexicon was broad but also specific enough to capture different emotions in text, Fitzgerald's novels contain more emotional content.
+The second result, aided by the NRC lexicon, does not provide conclusive results.
+Both authors cover a wide ground, and are tilted towards different emotions.
+However, this result appears skewed owing to the absence of adverbs.
+Fitzgerald's prose -- rich in metaphor, splattered with creative adverbs -- loses its sheen when placed under the microscope of this analysis.
+We are wary about accepting this result, and hope that a richer lexicon might capture what escaped us.
 
-As our analysis reveals, Hemingway is not emotionless by any means.
-Instead, his writing style is the primary reason because of which he appears to express less emotion in his writing.
-Thus, our hypothesis assessed emotion through a narrow lens and a better way of understanding it would be that Fitzgerald wrote with more easily identifiable emotion words than Hemingway, rather than assuming Hemingway’s characters lacked emotional depth.
-.
-They also helped to counter some initial fleeting assumptions we made about the corpus, like Fitzgerald’s characters being sadder or Hemingway’s characters being one-dimensionally angry, and allowed for a more nuanced understanding of their characters.
+The NER analysis combined with sentiment analysis, however, gives us a clearer picture.
+Fitzgerald's characters do appear to have emotional valences significantly higher than Hemingway's characters.
+While Hemingway's protagonists are more balanced in emotion than Fitzgerald's protagonists, it seems that this very characteristic of Fitzgerald's characters, their overt tilt towards specific emotions, are reflective of the high emotional density of his works.
+His verbosity in expression appears to contribute to detection of explicitly emotional content of his novels.
+
+Thus, we conclude that by using both a specific and general lexicon, accurately extracting entities, and assigning meaningful sentiment value to the characters, we have found Fitzgerald's works to be richer in emotion.
+With that said, better tools and a more fine-grained analysis is required to substantiate without doubt which author's work are richer in emotional content.
 
@@
 Another issue with our analysis was related to how we generated an overall emotion score, instead of understanding sentiment around a concept (such as sentiment around war in novels).
-This proved to be slightly broad, as compared to drilling down on a topic, however, it was appropriate for our hypothesis which sought to explore emotion in their writing styles.
+This proved to be slightly broad, as compared to drilling down on a topic, however, it was appropriate for our hypothesis which sought to explore emotion in the writing styles of the authors.
+
+On a more granular level, having the ability to wrangle data in endless ways proved to be a huge advantage.
+Testing various ideas by restructuring data in creative ways was a swift process.
+We were able to generate a multitude of visualisations by making very minor changes in code, which sped up the ideation and the subsequent verification process.
+One glaring flaw, though, was the NER process.
+In joining entities with sentences, the code relied on partial matches, which led to entities being attached to wrong sentences.
+And manually combing through and cleaning the data was a tedious and very time consuming task.
+Perhaps a better way could be devised to computationally attach entities to the right sentences.

### 2021-10-14 project_2.Rmd
Added final HTML. Project complete.

@@
   interactive = TRUE,
-  use.label = TRUE,
   size = 2,
+  grid.min = 1,
+  centre.y = 0,
   legend.position = "right"
 )
+data_vis_2
 ```

### 2021-10-14 Submission.Rmd
Fixing typos

@@
 
-The novels of F. Scott Fitzgerald contain more emotion than the novels of Earnest Hemingway.
+The novels of F. Scott Fitzgerald contain more emotion than the novels of Ernest Hemingway.
 
@@
 The corpus was created based on availability.
-Hemingway's works were not on available on Gutenberg, and had to be obtained from a different resource.
+Hemingway's works were not available on Gutenberg, and had to be obtained from a different resource.
 The files thus obtained were not clean, and contained textual artifacts.
@@
 One may argue that even simple descriptive paragraphs have emotion attached to them.
-But it is important to note that our tools are only equipped tell us which author’s work is more overtly emotional: that is, which author’s work will let us decipher emotion without having to analyse subtext or metaphors.
+But it is important to note that our tools are only equipped to tell us which author’s work is more overtly emotional: that is, which author’s work will let us decipher emotion without having to analyse subtext or metaphors.
 Close reading may or may not corroborate our findings, and lack of emotion in our findings definitely does not mean that an author employs less emotion in his writing.
@@
 
-As the graph indicates, Fitzgerald work has a higher percentage of these emotion words.
+As the graph indicates, Fitzgerald's work has a higher percentage of these emotion words.
 His least emotional novel, _The Great Gatsby_, with only _2.5%_ of the words matching our lexicon, is still higher than Hemingway's most emotional novel -- _A Farewell To Arms_, which stands at _2.1%_.

### 2021-10-14 Submission.Rmd
.

@@
 ```
-### Understanding Emotion in Hemingway's and Fitzgerald's work
+### Throwing Fitz in E(a)rnest: An Analysis of Emotion in Fitzgerald and Hemingway
 

### 2021-10-15 NER_analysis_final.Rmd
Negation in SA

@@
 
+```{r bigrams}
+nrc_pos <- get_sentiments("nrc")
+nrc_neg <- all_sentiments %>% mutate(negation = paste("not ", word)) %>% select(-word) %>% mutate(sentiment = paste("not", sentiment))
+
+nrc_neg <- rename(nrc_neg, bigram = negation) 
+nrc_neg <- nrc_neg %>% relocate(sentiment, .after = bigram)
+
+entities_bigram <- clean_entities %>%
+  unnest_tokens(bigram, sentence, token = "ngrams", n = 2)
+entities_bigram_sep <- entities_bigram %>% separate(bigram, c("word1", "word2"), sep = " ")
+
+entities_neg <- entities_bigram_sep %>% filter(word1 == "not") %>% inner_join(nrc_pos, by = c(word2 = "word"))
+
+entities_neg <- entities_neg %>% select(-word1) %>% rename(word = word2) %>% mutate(sentiment = paste("not", sentiment))
+
+entities_neg_dis <- entities_neg %>% distinct(sentence_num)
+
+which(entities_sentiment$sentence_num == entities_neg_dis)
+
+entities_neg <- entities_neg %>% select(-author, -title, -words, -kind, -X, -word1, -word2) %>% mutate(sentiment2 = paste("not", sentiment))
+entities_neg <- rename(entities_neg, sentiment_neg = sentiment)
+
+bigrams_separated <- entities_bigram %>%
+  separate(bigram, c("word1", "word2"), sep = " ")
+
+not_words <- bigrams_separated %>%
+  filter(word1 == "not") %>%
+  inner_join(nrc_pos, by = c(word2 = "word")) %>%
+  count(word2, sentiment, sort = TRUE)
+
+entities_bigram %>% count(bigram, sort = TRUE)
+
+entities_sentiment_bi <- entities_unnest %>%
+  group_by(author, title) %>%
+  inner_join(get_sentiments("nrc"))
+```
+
 ```{r message=FALSE}
@@
   group_by(author, title) %>%
-  inner_join(get_sentiments("nrc")) %>%
+  inner_join(get_sentiments("nrc"))
+entities_sentiment <- bind_rows(entities_sentiment, entities_neg) %>%
 count(sentence_num, sentiment) %>%
-  spread(sentiment, n, fill = 0) %>%
-  mutate(sentiment = positive - negative)
+  spread(sentiment, n, fill = 0)
+  entities_sentiment[ ,4] = 0 - entities_sentiment[ ,4]
+entities_sentiment[ ,6] = 0 - entities_sentiment[ ,6]
+entities_sentiment[ ,7] = 0 - entities_sentiment[ ,7]
+entities_sentiment[ ,10] = 0 - entities_sentiment[ ,10]
+entities_sentiment[ ,11] = 0 - entities_sentiment[ ,11]
+entities_sentiment[ ,18] = 0 - entities_sentiment[ ,18]
+entities_sentiment[ ,19] = 0 - entities_sentiment[ ,19]
+entities_sentiment[ ,21] = 0 - entities_sentiment[ ,21]
+  entities_sentiment <- entities_sentiment %>% select(-sentence_num) %>% mutate(sentiment = rowSums(across(where(is.numeric))))
+entities_sentiment <- entities_sentiment %>% select(title, author, sentiment)
 ```
@@
   geom_col()  +
-  facet_wrap(~ title, scales = "free", ncol = 2) +coord_flip() + labs(x = "Names of characters", 
+  facet_wrap(~ title, scales = "free") +coord_flip() + labs(x = "Names of characters", 
        y = "% of positive emotion", fill = "Novel", title = "% of Positive Emotions Detected")
@@
   geom_col() +
-  facet_wrap(~ title, scales = "free") + labs(x = "Names of characters", 
-       y = "% of negative emotion", fill = "Novel", title = "% of Negative Emotions Detected")
+  facet_wrap(~ title, scales = "free") + labs(x = "% of Negative Emotion", 
+       y = "Characters in Novels", fill = "Novel", title = "% of Negative Emotions Detected")
 vis_ner_2_int <- ggplotly(vis_ner_2, tooltip = "text")

### 2021-10-15 project_2.Rmd
.

@@
 #table shows each text and the respective percentage
-fitzgerald_emotion_calc <- fitzgerald_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = n/sum(n)*100) %>% mutate(author = "F. Scott Fitzgerald")
+fitzgerald_emotion_calc <- fitzgerald_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = round(n/sum(n)*100, 2)) %>% mutate(author = "F. Scott Fitzgerald")
 
@@
 #table shows each text and the respective percentage
-hemingway_emotion_calc <- hemingway_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = n/sum(n)*100) %>% mutate(author = "Earnest Hemingway")
+hemingway_emotion_calc <- hemingway_emotions %>% group_by(title) %>% count(emotion) %>% mutate(percent = round(n/sum(n)*100, 2)) %>% mutate(author = "Earnest Hemingway")
 
